### 2023-03-21 stats/analysis.Rmd

---
title: "Stats"
author: "Julia Prein"
date: "03/21/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse) # data handling etc.
library(viridis) # col scales, plotting...
library(ggpubr)
library(ggthemes)
library(tidyboot) # for bootstrap CIs
library(brms) # modeling
library(tidybayes) # for stat_halfeye

options(scipen = 999)
theme_set(theme_classic())
# Seed for random number generation
set.seed(42)
```

```{r load_data}
# testtrials <- readRDS(file = "../data/tango-testtrials.rds") %>% 
#   # determine order of factors
#   mutate(
#     targetPosition = factor(targetPosition, levels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "box1", "box2", "box3", "box4", "box5")), 
#     studyversion = factor(studyversion, levels = c("hedge", "box")), 
#     datacollection = factor(datacollection, levels = c("in-person - supervised", "remote - unsupervised")),
#     sample = factor(sample, levels = c("kids", "adults")), 
#     studytype = factor(studytype, levels = c("vali", "vali2", "reli")),
#     
#     # for models, summarize targetPositions in a way that it's symmetrical for left/ right side
#     # the smaller the value, the more central the position
#     symmetricPosition = ifelse(studyversion == "box", 
#                                # for box version
#                                case_when(
#                                   targetPosition == "box3" ~ 0, 
#                                   targetPosition == "box2" | targetPosition == "box4" ~ 1,
#                                   targetPosition == "box1" | targetPosition == "box5" ~ 2), 
#                                # for hedge version
#                                targetCentralityX
#                        ), 
#     
#     # distance between target and click in balloon widths (balloon width is 160 SVG units)
#     clickDistInBalloons = abs(clickDistFromTargetCenterX)/160,
#   )
```