### 2023-03-21 supplements/supplements.Rmd

---
title             : "Supplements for the manuscript  'Variation in gaze understanding across the life span: A process-level perspective'"
shorttitle        : "Supplements"
author:

  - name          : "Julia Prein"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Max Planck Institute for Evolutionary Anthropology, Deutscher Platz 6, 04103 Leipzig, Germany"
    email         : "julia_prein@eva.mpg.de"

  - name          : "Manuel Bohn"
    affiliation   : "1"
  - name          : "Luke Maurits"
    affiliation   : "1"
  - name          : "Annika Werwach"
    affiliation   : "1"
  - name          : "Daniel B. M. Haun"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "Department of Comparative Cultural Psychology, Max Planck Institute for Evolutionary Anthropology, Leipzig, Germany"
floatsintext      : yes
bibliography      : ["../../references.bib"]
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
documentclass     : "apa7"
classoption       : "man"
output            : papaja::apa6_word
header-includes:
  - \usepackage{setspace}
  - \captionsetup[figure]{font={stretch=1}}
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include = FALSE}
library("papaja")
library("tidyverse")
library("ggpubr")
library("kableExtra")
theme_set(theme_classic())
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

\newpage

# References

```{=tex}
\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
```
::: {#refs custom-style="Bibliography"}
:::

```{=tex}
\endgroup
```