### 2023-03-22 stats/analysis.Rmd
add already written notes/code

---
title: "Stats"
author: "Julia Prein"
date: "03/21/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse) # data handling etc.
library(viridis) # col scales, plotting...
library(ggpubr)
library(ggthemes)
library(tidyboot) # for bootstrap CIs
library(brms) # modeling
library(tidybayes) # for stat_halfeye

options(scipen = 999)
theme_set(theme_classic())
# Seed for random number generation
set.seed(42)
```

```{r load_data}
testtrials <- readRDS(file = "../data/tango-testtrials.rds") %>%
  mutate_if(is.character, as.factor) %>% 
  # determine order of factors
  mutate(
    targetPosition = factor(targetPosition, levels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "box1", "box2", "box3", "box4", "box5")),
    studyversion = factor(studyversion, levels = c("hedge", "box")),
    datacollection = factor(datacollection, levels = c("in-person - supervised", "remote - unsupervised")),
    sample = factor(sample, levels = c("kids", "teens", "adults")),
    studytype = factor(studytype, levels = c("vali", "vali2", "reli")),

    # for models, summarize targetPositions in a way that it's symmetrical for left/ right side
    # the smaller the value, the more central the position
    symmetricPosition = case_when(
                                  targetPosition == "5" | targetPosition == "6" | targetPosition == "box3" ~ 1,
                                  targetPosition == "4" | targetPosition == "7" | targetPosition == "box2" | targetPosition == "box4" ~ 2,
                                  targetPosition == "3" | targetPosition == "8" | targetPosition == "box1" | targetPosition == "box5" ~ 3,
                                  targetPosition == "2" | targetPosition == "9" ~ 4,
                                  targetPosition == "1" | targetPosition == "10" ~ 5,
                       ),

    # distance between target and click in balloon widths (balloon width is 160 SVG units)
    clickDistInBalloons = abs(clickDistFromTargetCenterX)/160,
  )
```

```{r check_sample}
# child sample 
testtrials %>% 
  filter(agegroup < 18) %>% 
  group_by(datacollection, studytype, studyversion, agegroup) %>%
  mutate(female = ifelse(gender == "f", 1, 0)) %>% 
  summarise(
    nTotal = n_distinct(subjID), 
    meanAge = mean(age, na.rm = T) %>% round(2),
    sdAge = sd(age, na.rm = T) %>% round(2),
    minAge = min(age, na.rm = T),
    maxAge = max(age, na.rm = T),
    nFemales = ceiling(sum(female/15)),
  ) %>% View()

# reli
testtrials %>% 
  filter(agegroup < 6 & completedReli) %>% 
  group_by(datacollection, studyversion, agegroup) %>%
  mutate(female = ifelse(gender == "f", 1, 0)) %>% 
  summarise(
    nTotal = n_distinct(subjID), 
    meanAge = mean(age, na.rm = T) %>% round(2),
    sdAge = sd(age, na.rm = T) %>% round(2),
    minAge = min(age, na.rm = T),
    maxAge = max(age, na.rm = T),
    nFemales = ceiling(sum(female/15)),
  )
```

# DEVELOPMENTAL TRAJECTORY

```{r variation_preparateData}
variationInd <- testtrials %>% 
  mutate(
    age_centered = scale(age, center = TRUE, scale = TRUE)[,1],
  ) %>% 
  group_by(studyversion, sample, datacollection, age, agegroup, age_centered, subjID) %>%
  summarise(mean = mean(clickDistInBalloons, na.rm = TRUE))

# variation <- variationInd %>%
#   group_by(sample, datacollection, ageInYears) %>%
#   tidyboot_mean(column = mean)
```

```{r dev_model}
mDataAgeHedge <- testtrials %>% 
  filter(studyversion == "hedge" & studytype == "vali") %>% 
  mutate(
    # scaled (centered around mean, SD = 1, i.e., z-transformed)
    age = scale(age, center = TRUE, scale = TRUE)[,1], 
  ) %>% 
  select(subjID, age, agegroup, datacollection, symmetricPosition, targetPosition, clickDistInBalloons)

mAgeHedge <- brm(clickDistInBalloons ~ age + datacollection + (symmetricPosition | subjID) + (1 | targetPosition),
                    data = mDataAgeHedge,
                    family = lognormal(),
                    warmup = 500, 
                    iter   = 1000, # todo: increase once done
                    chains = 4, 
                    inits  = "random",
                    cores  = 100)

write.csv(mAgeHedge, "../saves/mAgeHedge.csv", quote = F, row.names = F)
saveRDS(mAgeHedge, "../saves/mAgeHedge.rds")
```

```{r dev_plot}
ggplot(data = variationInd, aes(x = age, y = mean, shape = datacollection, color = studyversion)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("#006c66", "darkgrey")) +
  geom_smooth(data = variationInd, aes(x = age, y = mean, group = studyversion, color = studyversion), method = "lm", size = 1, se=F) +
  facet_wrap(. ~ sample, scales = "free") +
  scale_y_continuous(limits = c(0, 6), breaks = seq(0, 6, 1))

ggsave("../figures/tango_dev.png", width = 8, height = 4)

ggplot(data = variationInd, aes(x = age, y = mean, shape = datacollection, color = studyversion)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("#006c66", "darkgrey")) +
  geom_smooth(data = variationInd, aes(x = age, y = mean, group = studyversion, color = studyversion), size = 1, se=F) +
  # facet_wrap(. ~ sample, scales = "free") +
  scale_y_continuous(limits = c(0, 6), breaks = seq(0, 6, 1))
ggsave("../figures/tango_dev_wofacets.png", width = 8, height = 4)
```

# COGNITIVE MODEL

```{r gazefunnel_hedge}
imprecisionBinHedge <- testtrials %>% 
  filter(agegroup < 6 & studyversion == "hedge") %>% 
  group_by(datacollection, targetPosition) %>% 
  tidyboot_mean(col = clickDistInBalloons, na.rm = TRUE)

imprecisionBinHedgeInd <- testtrials %>% 
  filter(agegroup < 6 & studyversion == "hedge") %>% 
  group_by(datacollection, targetPosition, subjID) %>% 
  tidyboot_mean(col = clickDistInBalloons, na.rm = TRUE)

ggplot() +
  labs(x = "Target position (binned)", y = "Average imprecision in target widths", col = "Data collection mode") +
  # geom_point(data = imprecisionBinHedgeInd, aes(x = targetPosition, y = mean, col = datacollection), alpha = .1, position = position_dodge(width = 0.2)) +
  geom_pointrange(data = imprecisionBinHedge, aes(x = targetPosition, y = mean, ymin = ci_lower, ymax = ci_upper, col = datacollection), pch = 1,  size = 0.7, position = position_dodge(width = .2)) +
  scale_color_colorblind()

```