### 2023-06-05 stats/magnet.Rmd

---
title: "Stats"
author: "Julia Prein"
date: "03/21/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse) # data handling etc.
library(corrr)
library(psych)
library(apaTables)

options(scipen = 999)
theme_set(theme_classic())

# Seed for random number generation
set.seed(42)
```

```{r load_data}
magnet <- readRDS("../data/magnet-data.rds")
```

```{r scatterplots}
magnet %>% 
  ggplot(., aes(x = magnet_mean, y = tango_mean)) + 
  geom_abline(intercept = 0, slope = 1, col = "darkgrey", linetype = "longdash") +
  geom_smooth(method = "lm", size = 1, alpha = 0.25, col = "darkgrey") +
  geom_point(color = "#005555") +
  stat_cor(method = "pearson", label.x = 5, label.y = 6, aes(label = paste(..r.label..)), size = 5, r.accuracy = 0.01, cor.coef.name = "r") + 
  scale_x_continuous(limits = c(0, 6), breaks = seq(0, 6, 0.5)) +
  scale_y_continuous(limits = c(0, 6), breaks = seq(0, 6, 0.5)) +
  labs(x = "Magnet mean", y = "Tango mean")

magnet %>% 
  ggplot(., aes(x = magnet_mean, y = tango_mean)) + 
  geom_abline(intercept = 0, slope = 1, col = "darkgrey", linetype = "longdash") +
  geom_smooth(method = "lm", size = 1, alpha = 0.25, col = "darkgrey") +
  geom_point(color = "#005555") +
  stat_cor(method = "pearson", label.x = 3.5, label.y = 3.25, aes(label = paste(..r.label..)), size = 5, r.accuracy = 0.01, cor.coef.name = "r") + 
  scale_x_continuous(limits = c(0, 3.75), breaks = seq(0, 4, 0.5)) +
  scale_y_continuous(limits = c(0, 3.75), breaks = seq(0, 4, 0.5)) +
  labs(x = "Magnet mean", y = "Tango mean")
```

```{r corrplot}
magnet %>% 
  select(tango_mean, magnet_mean, tom_aggregate, perspective_aggregate) %>% 
  rename("Tango mean" = tango_mean, "Magnet mean" = magnet_mean, "ToM aggregate score" = tom_aggregate, "Perspective-taking score" = perspective_aggregate) %>% 
  cor() %>% 
  corrplot(diag = FALSE, type = 'upper', method = 'ellipse', order = 'FPC', col = COL2('BrBG'), tl.col = 'black', cl.ratio = 0.2, tl.srt = 30, addCoef.col ='black', number.cex = 0.8)
```

```{r cor_matrix}
tango_magnet <- magnet %>% select(tango_mean, magnet_mean) %>% corr.test(method = "pearson")
tango_tom <- magnet %>% select(tango_mean, tom_aggregate, perspective_aggregate) %>% corr.test(method = "spearman")
magnet_tom <- magnet %>% select(magnet_mean, tom_aggregate, perspective_aggregate) %>% corr.test(method = "spearman")

cor_matrix <- rbind(tango_magnet$ci, tango_tom$ci, magnet_tom$ci) %>% 
  mutate(Correlation = rownames(.)) %>%
  as.tibble()

cor_matrix <- cor_matrix %>%
  # remove tom - perspective correlations
  filter(Correlation != "tm_gg-prsp_1") %>%
  mutate(
    Correlation = recode(Correlation,
                  `tng_m-mgnt_` = "Tango x Magnet", 
                  `tng_m-tm_gg` = "Tango x ToM",
                  `tng_m-prsp_` = "Tango x Perspective-taking",
                  `mgnt_-tm_gg` = "Magnet x ToM",
                  `mgnt_-prsp_` = "Magnet x Perspective-taking", 
                  `tm_gg-prsp_` = "ToM x Perspective-taking")
  ) %>% 
  mutate(across(where(is.numeric), round, 2)) %>% 
  select(Correlation, r, lower, upper, p)

# apa table function only supports pearson corr so far
magnet %>% select(tango_mean, magnet_mean, tom_aggregate, perspective_aggregate) %>% apa.cor.table(filename = "APA_Correlation_Table.doc", table.number=1)
```

```{r plot_cor_ci}
cor_matrix %>% 
  filter(Correlation != "ToM x Perspective-taking") %>% 
  mutate(Correlation = fct_relevel(Correlation, "Magnet x Perspective-taking", "Magnet x ToM", "Tango x Perspective-taking", "Tango x ToM", "Tango x Magnet")) %>%
  ggplot(., aes(x=Correlation, y=r, ymin=lower, ymax=upper)) +
  geom_pointrange(color = "#005555", size = 0.8, linewidth = 1) +
  scale_y_continuous(limits = c(-1, 1), breaks = seq(-1, 1, 0.25)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab("") +
  coord_flip() +
  theme(axis.title.x = element_text(face="italic"))

ggsave("../figures/magnet_cor_ci.png", width = 5, height = 3)
```
```