### 2023-06-09 stats/gazemodel.Rmd

---
title: "Gaze model analysis"
author: "Julia Prein"
date: "06/07/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse) # data handling etc.
library(tidyboot) # for bootstrap CIs

options(scipen = 999)
theme_set(theme_classic())

# Seed for random number generation
set.seed(42)
```

```{r load_data}
gaze_data_kids <- readRDS("../data/tango-model-kids.rds")
gaze_data_adults <- readRDS("../data/tango-model-adults.rds") %>% mutate(age_group = 18)
gaze_data <- full_join(gaze_data_kids, gaze_data_adults)
```

```{r sample}
gazemodel_sample_kids <- gaze_data_kids %>% 
  group_by(age_group) %>%
  mutate(female = ifelse(gender == "f", 1, 0)) %>% 
  summarise(
    n_total = n_distinct(subj_id), 
    age_mean = mean(age, na.rm = T) %>% round(2),
    age_sd = sd(age, na.rm = T) %>% round(2),
    age_min = min(age, na.rm = T),
    age_max = max(age, na.rm = T),
    n_females = ceiling(sum(female/15)),
  )

gazemodel_sample_kids

gazemodel_sample_adults <- gaze_data_adults %>% 
  mutate(female = ifelse(gender == "f", 1, 0)) %>% 
  summarise(
    n_total = n_distinct(subj_id), 
    age_mean = mean(age, na.rm = T) %>% round(2),
    age_sd = sd(age, na.rm = T) %>% round(2),
    age_min = min(age, na.rm = T),
    age_max = max(age, na.rm = T),
    n_females = ceiling(sum(female/15)),
  ) %>% 
  mutate(age_group = 18)

gazemodel_sample <- gazemodel_sample_kids %>% full_join(gazemodel_sample_adults)

write.csv(gazemodel_sample, "../saves/gazemodel_sample.csv", quote = F, row.names = F)
saveRDS(gazemodel_sample, "../saves/gazemodel_sample.rds")
```

```{r plot}
dev_plot <- gaze_data_kids %>% 
  group_by(subj_id, age, age_group) %>%
  summarise(mean = mean(click_dist_balloons, na.rm = TRUE)) %>% 
  ggplot(., aes(x = age, y = mean)) +
  geom_point(color = "darkgrey", alpha = 0.5) +
  geom_smooth(method = "lm", size = 1, se = T, color = "#006c66", fill = "#006c66") + 
  scale_x_continuous(limits = c(3, 5), breaks = seq(3, 5, 0.5)) +
  scale_y_continuous(limits = c(0, 5), breaks = seq(0, 5, 1)) +
  labs(x = "Age in years", y = "Average imprecision in target widths")

dev_plot

# ggsave("../figures/dev_plot.png", width = 8, height = 4, scale = 1.2)
```

```{r gazefunnel}
target_position_imprecision <- gaze_data %>%
  group_by(age_group, target_position) %>%
  tidyboot_mean(column = click_dist_balloons, na.rm = TRUE)

target_position_imprecision_ind <- gaze_data %>%
  group_by(subj_id, age_group, target_position) %>%
  summarize(mean = mean(click_dist_balloons, na.rm = TRUE))

gazefunnel <- ggplot() +
  geom_point(data = target_position_imprecision_ind, aes(x = target_position, y = mean, col = as.factor(age_group)), alpha = .1, position = position_dodge(width = 0.3)) +

  geom_pointrange(data = target_position_imprecision, aes(x = target_position, y = mean, ymin = ci_lower, ymax = ci_upper, col = as.factor(age_group)), pch = 18,  size = 1, position = position_dodge(width = .3)) +

  scale_y_continuous(limits = c(0, 7), breaks = seq(0, 7, 1)) +
  scale_color_manual(values = c("#c6d325", "#00b1ea", "#29485d", "darkgrey"), breaks = c("3", "4", "5", "18"), labels = c("3-year-olds", "4-year-olds", "5-year-olds", "Adults")) +
    labs(x = "Target position (binned)", y = "Average imprecision in target widths", col = "Age in years")

gazefunnel

ggsave("../figures/gazefunnel.png", width = 6, height = 4, scale = 1.2)
saveRDS(gazefunnel, "../saves/gazefunnel.rds")
```