### 2023-06-15 stats/supplementary_analyses.Rmd
update magnet data (remove safety copies), add suppl. human vs animal stimuli analysis

---
title: "Supplementary analysis"
author: "Julia Prein"
date: "06/13/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse) # data handling etc.
library(tidybayes) # for stat_halfeye, add_predictive_draws
library(janitor)
library(brms) # modeling
library(cmdstanr) # modeling
check_cmdstan_toolchain(fix = TRUE, quiet = TRUE)

options(scipen = 999)
theme_set(theme_classic())

# Seed for random number generation
set.seed(42)
```

```{r load_data}
tango_humans <- readRDS("../data/magnet-tango-data.rds") %>% 
  mutate(target_position = as.factor(target_position), stimuli = "humans")

tango_animals <- readRDS("../data/tango-lifespan.rds") %>%
  filter(age >= min(tango_humans$age) & age <= max(tango_humans$age)) %>% 
  mutate(stimuli = "animals")

combined_data <- full_join(tango_humans, tango_animals) %>% 
  remove_empty(which = "cols") %>% 
  mutate(
    age_centered = age - min(age)
  )
```

```{r exploratory_plots}
combined_data %>% 
  ggplot(., aes(x = agent, y = click_dist_balloons)) +
  geom_count()
```

```{r models, message=FALSE, warning=FALSE}
mdata <- combined_data %>% 
  mutate(
    # scaled (centered around mean, SD = 1, i.e., z-transformed)
    age_scaled = scale(age), 
    symmetric_position = scale(symmetric_position),
    trial_nr = trial_nr - min(trial_nr), 
    click = abs(click_dist_from_target_center_x), 
  ) %>% 
  select(subj_id, age, age_scaled, stimuli, agent, symmetric_position, trial_nr, click) %>% 
  filter(click > 0)

mstimuli <- brm(click ~ age_scaled + stimuli + symmetric_position + trial_nr + (1 + symmetric_position + trial_nr | subj_id),
                    data = mdata,
                    family = lognormal(),
                    warmup = 2000, 
                    iter   = 6000, 
                    chains = 4, 
                    cores  = 4, 
                    threads = threading(8), 
                    control = list(adapt_delta = 0.95), 
                    backend = "cmdstanr") %>% add_criterion(c("loo","waic"))

mstimuli

pp_check(mstimuli)

magent <- brm(click ~ age_scaled + agent + symmetric_position + trial_nr + (1 + symmetric_position + trial_nr | subj_id),
                    data = mdata,
                    family = lognormal(),
                    warmup = 2000, 
                    iter   = 6000, 
                    chains = 4, 
                    cores  = 4, 
                    threads = threading(8), 
                    control = list(adapt_delta = 0.95), 
                    backend = "cmdstanr") %>% add_criterion(c("loo","waic"))

magent

pp_check(magent)

saveRDS(mstimuli, "../saves/supplements_magnet_mstimuli.rds")
saveRDS(magent, "../saves/supplements_magnet_magent.rds")
```