### 2023-07-07 stats/lifespan.Rmd

---
title: "Lifespan analysis"
author: "Julia Prein"
date: "03/21/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse) # data handling etc.
library(tidybayes) # for stat_halfeye, add_predictive_draws
library(tidyboot) # for bootstrapped CIs
library(brms) # modeling
library(cmdstanr) # modeling
check_cmdstan_toolchain(fix = TRUE, quiet = TRUE)

options(scipen = 999)
theme_set(theme_classic())

# Seed for random number generation
set.seed(42)
```

```{r load_data}
lifespan_data <- readRDS("../data/tango-lifespan.rds")
```

```{r prepare_modeldata}
mdata <- lifespan_data %>% 
  group_by(subj_id, age, age_centered) %>% 
  summarize(mean_dist = mean(click_dist_balloons, na.rm = TRUE))
```

```{r sample}
lifespan_sample <- lifespan_data %>% 
  group_by(age_group) %>%
  mutate(female = ifelse(gender == "f", 1, 0)) %>% 
  summarise(
    n_total = n_distinct(subj_id), 
    age_mean = mean(age, na.rm = T) %>% round(2),
    age_sd = sd(age, na.rm = T) %>% round(2),
    age_min = min(age, na.rm = T),
    age_max = max(age, na.rm = T),
    n_females = ceiling(sum(female/15)),
  )

lifespan_sample

write.csv(lifespan_sample, "../saves/lifespan_sample.csv", quote = F, row.names = F)
saveRDS(lifespan_sample, "../saves/lifespan_sample.rds")
```

```{r plot}
lifespan_plot <- lifespan_data %>% 
  group_by(subj_id, age, age_group) %>%
  summarise(mean = mean(click_dist_balloons, na.rm = TRUE)) %>% 
  ggplot(., aes(x = age, y = mean)) +
  geom_point(color = "darkgrey", alpha = 0.5) +
  geom_smooth(method = "loess", size = 1, se = T, color = "#006c66", fill = "#006c66") + 
  scale_x_continuous(limits = c(0, 80), breaks = seq(0, 80, 5)) + 
  scale_y_continuous(limits = c(0, 4), breaks = seq(0, 4, 0.5)) + 
  labs(x = "Age in years", y = "Average imprecision in target widths")

lifespan_plot

ggsave("../figures/lifespan_plot.png", width = 8, height = 4, scale = 1.2)
saveRDS(lifespan_plot, "../saves/lifespan_plot.rds")
```

```{r models, message=FALSE, warning=FALSE}
# linear age
mlinear <- brm(mean_dist ~ age_centered,
                    data = mdata,
                    family = lognormal(),
                    warmup = 2000, 
                    iter   = 6000, 
                    chains = 4, 
                    cores  = 4, 
                    threads = threading(8), 
                    control = list(adapt_delta = 0.95), 
                    backend = "cmdstanr") %>% add_criterion(c("loo","waic"))

mlinear

pp_check(mlinear)

write.csv(mlinear, "../saves/lifespan_mlinear.csv", quote = F, row.names = F)
saveRDS(mlinear, "../saves/lifespan_mlinear.rds")

## quadratic age
mquadratic <- brm(mean_dist ~ 1 + age_centered + I(age_centered^2),
                    data = mdata,
                    family = lognormal(),
                    warmup = 2000, 
                    iter   = 6000, 
                    chains = 4, 
                    cores  = 4, 
                    threads = threading(8), 
                    control = list(adapt_delta = 0.95), 
                    backend = "cmdstanr") %>% add_criterion(c("loo","waic"))

mquadratic

pp_check(mquadratic)

conditional_effects(mquadratic, ndraws = 200, spaghetti = TRUE) %>% 
  plot(points = T)

write.csv(mquadratic, "../saves/lifespan_mquadratic.csv", quote = F, row.names = F)
saveRDS(mquadratic, "../saves/lifespan_mquadratic.rds")

## cubic age
mcubic <- brm(mean_dist ~ 1 + age_centered + I(age_centered^2) + I(age_centered^3),
                    data = mdata,
                    family = lognormal(),
                    warmup = 2000, 
                    iter   = 6000, 
                    chains = 4, 
                    cores  = 4, 
                    threads = threading(8), 
                    control = list(adapt_delta = 0.95), 
                    backend = "cmdstanr") %>% add_criterion(c("loo","waic"))

mcubic

pp_check(mcubic)

plot(mcubic)

conditional_effects(mcubic, ndraws = 200, spaghetti = TRUE) %>% 
  plot(points = T)

write.csv(mcubic, "../saves/lifespan_mcubic.csv", quote = F, row.names = F)
saveRDS(mcubic, "../saves/lifespan_mcubic.rds")

## gaussian process model
mgaussianprocess <- brm(mean_dist ~ gp(age_centered, k=50, c=5/4, scale=TRUE),
                    data = mdata,
                    family = lognormal(),
                    warmup = 2000, 
                    iter   = 6000, 
                    chains = 4, 
                    cores  = 4, 
                    threads = threading(8), 
                    control = list(adapt_delta = 0.95), 
                    backend = "cmdstanr") %>% add_criterion(c("loo","waic"))

mgaussianprocess

pp_check(mgaussianprocess)

conditional_effects(mgaussianprocess, ndraws = 200, spaghetti = TRUE) %>% 
  plot(points = T)

write.csv(mgaussianprocess, "../saves/lifespan_mgaussianprocess.csv", quote = F, row.names = F)
saveRDS(mgaussianprocess, "../saves/lifespan_mgaussianprocess.rds")
```

```{r lifespan_modelcomparison}
# mlinear <- readRDS("../saves/lifespan_mlinear.rds")
# mquadratic <- readRDS("../saves/lifespan_mquadratic.rds")
# mcubic <- readRDS("../saves/lifespan_mcubic.rds")
# mgaussianprocess <- readRDS("../saves/lifespan_mgaussianprocess.rds")

mcomparison <- loo_compare(mlinear, mquadratic, mcubic, mgaussianprocess, criterion = "waic") %>% 
  as_tibble(rownames = "model") %>%
  left_join(as_tibble(model_weights(mlinear, mquadratic, mcubic, mgaussianprocess, weights = "waic"), rownames = "model")) %>% 
  rename(modelweight = value)

mcomparison %>% 
  select(-c(se_elpd_waic, p_waic, se_p_waic)) %>% 
  mutate(across(where(is.numeric), round, 4))

write.csv(mcomparison, "../saves/lifespan_mcomparison.csv", quote = F, row.names = F)
saveRDS(mcomparison, "../saves/lifespan_mcomparison.rds")
```

```{r draws_winning_model}
# mgaussianprocess <- readRDS("../saves/lifespan_mgaussianprocess.rds")

mdraws <- mdata %>%
  add_predicted_draws(mgaussianprocess) %>% 
  ungroup() %>% 
  select(-c(.row, .chain, .iteration)) %>%
  rename(draw = .draw, predicted_value = .prediction)

write.csv(mdraws, "../saves/lifespan_mdraws.csv", quote = F, row.names = F)
saveRDS(mdraws, "../saves/lifespan_mdraws.rds")
```

```{r prediction_plot_winning_model}
### this sections takes a while to run

# mdraws <- readRDS("../saves/lifespan_mdraws.rds")

mdraws %>%
  ggplot(aes(x = age)) +
  stat_lineribbon(aes(y = predicted_value), .width = c(.95), colour = "#006c66", fill = "lightgrey") +
  geom_jitter(aes(y = mean_dist), colour = "darkgrey", size = 1.5, alpha = 0.8) +
  scale_x_continuous(limits = c(0,80), breaks = seq(0, 80, 5)) +
  scale_y_continuous(limits = c(0,4)) +
  labs(x = "Age in years", y = "Mean click distance in target widths")

# saveRDS(prediction_plot, file = "../saves/lifespan_modelplot.rds")
ggsave("../figures/lifespan_modelplot.png", width = 8, height = 5, scale = 1, bg = "white")
```

```{r playground}
# mgaussianprocess <- readRDS("../saves/lifespan_mgaussianprocess.rds")

# model checks: 
# density plot overlay of the original response values (the thick black line) with 10 draws from the posterior distribution of the model
pp_check(mgaussianprocess)

# empirical cumulative distribution function of the observations and random draws from the model posterior:
pp_check(mgaussianprocess, type = "ecdf_overlay")

# plots distribution of estimates & chains converging
plot(mgaussianprocess)

# model plotting
spline_plot_data <- fitted(mgaussianprocess) %>% 
  data.frame() %>% 
  bind_cols(mdata)

spline_plot_data %>% 
  ggplot(aes(x = age, y = mean_dist, ymin = Q2.5, ymax = Q97.5)) + 
  geom_hline(yintercept = fixef(mgaussianprocess)[1, 1], linetype = 2) +
  geom_point(color = "steelblue") +
  geom_ribbon(alpha = 2/3) +
  labs(x = "Age",
       y = "Imprecision") +
  theme_classic()

# predictions & actual raw data underneath
conditional_effects(mgaussianprocess, ndraws = 200, spaghetti = TRUE) %>% 
  plot(points = T)

# get intercepts
posterior_summary(mgaussianprocess)
```

```{r smooths}
lifespan_smooths <- mdata %>%
  ggplot(aes(x = age, y = mean_dist)) +  
  geom_jitter(colour = "grey", size = 1.5, alpha = 0.8) + 
  geom_smooth(method = lm, se = F, aes(colour = "A")) +
  geom_smooth(method = lm, formula = y ~ poly(x, 2), se = F, aes(colour = "B")) +
  geom_smooth(method = lm, formula = y ~ poly(x, 3), se = F, aes(colour = "C")) +
  geom_smooth(method = loess, se = F, aes(colour = "D")) +
  geom_smooth(method = lm, formula = y ~ splines::bs(x, 5), se = F, aes(colour = "E")) +
  geom_smooth(method = gam, formula = y ~ s(x), se = F, aes(colour = "F")) +
  scale_fill_brewer(palette = "Greys") + 
  scale_x_continuous(limits = c(0,80), breaks = seq(0, 80, 5)) + 
  scale_y_continuous(limits = c(0,4)) + 
  scale_color_manual(name = "Smooth method", 
                     breaks = c("A", "B", "C", "D", "E", "F"),
                     values = c("#c6d325", "#00b1ea", "#29485d", "#9D3BBD", "#F71C8D", "orange"),
                     labels = c("linear", "quadratic", "cubic", "loess", "splines (k=5)", "splines (default)")) +
  labs(x = "Age in years", y = "Mean click distance in target widths")

lifespan_smooths

saveRDS(lifespan_smooths, file = "../saves/lifespan_smooths.rds")
ggsave("../figures/lifespan_smooths.png", width = 10, height = 6, scale = 0.7, bg = "white")
```