### 2023-07-10 stats/lifespan.Rmd
clean up, add Bayesian changepoint (Rbeast)

---
title: "Lifespan analysis"
author: "Julia Prein"
date: "03/21/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse) # data handling etc.
library(tidybayes) # for stat_halfeye, add_predictive_draws
library(tidyboot) # for bootstrapped CIs
library(brms) # modeling
library(cmdstanr) # modeling
library(Rbeast) # Bayesian change point analysis
check_cmdstan_toolchain(fix = TRUE, quiet = TRUE)

options(scipen = 999)
theme_set(theme_classic())

# Seed for random number generation
set.seed(42)
```

```{r load_data}
lifespan_data <- readRDS("../data/tango-lifespan.rds")
```

```{r prepare_modeldata}
mdata <- lifespan_data %>% 
  group_by(subj_id, age, age_centered) %>% 
  summarize(mean_imprecision = mean(click_dist_balloons, na.rm = TRUE))
```

```{r sample}
lifespan_sample <- lifespan_data %>% 
  group_by(age_group) %>%
  mutate(female = ifelse(gender == "f", 1, 0)) %>% 
  summarise(
    n_total = n_distinct(subj_id), 
    age_mean = mean(age, na.rm = T) %>% round(2),
    age_sd = sd(age, na.rm = T) %>% round(2),
    age_min = min(age, na.rm = T),
    age_max = max(age, na.rm = T),
    n_females = ceiling(sum(female/15)),
  )

lifespan_sample

write.csv(lifespan_sample, "../saves/lifespan_sample.csv", quote = F, row.names = F)
saveRDS(lifespan_sample, "../saves/lifespan_sample.rds")
```

```{r plot}
lifespan_plot <- lifespan_data %>% 
  group_by(subj_id, age, age_group) %>%
  summarise(mean = mean(click_dist_balloons, na.rm = TRUE)) %>% 
  ggplot(., aes(x = age, y = mean)) +
  geom_point(color = "darkgrey", alpha = 0.5) +
  geom_smooth(method = "loess", size = 1, se = T, color = "#006c66", fill = "#006c66") + 
  scale_x_continuous(limits = c(0, 80), breaks = seq(0, 80, 5)) + 
  scale_y_continuous(limits = c(0, 4), breaks = seq(0, 4, 0.5)) + 
  labs(x = "Age in years", y = "Average imprecision in target widths")

lifespan_plot

ggsave("../figures/lifespan_plot.png", width = 8, height = 4, scale = 1.2)
saveRDS(lifespan_plot, "../saves/lifespan_plot.rds")
```

# MODEL COMPARISON FOR DEVELOPMENTAL TRAJECTORY

```{r models, message=FALSE, warning=FALSE}
# linear age
mlinear <- brm(mean_imprecision ~ age_centered,
                    data = mdata,
                    family = lognormal(),
                    warmup = 2000, 
                    iter   = 6000, 
                    chains = 4, 
                    cores  = 4, 
                    threads = threading(8), 
                    control = list(adapt_delta = 0.95), 
                    backend = "cmdstanr") %>% add_criterion(c("loo","waic"))

mlinear

pp_check(mlinear)

write.csv(mlinear, "../saves/lifespan_mlinear.csv", quote = F, row.names = F)
saveRDS(mlinear, "../saves/lifespan_mlinear.rds")

## quadratic age
mquadratic <- brm(mean_imprecision ~ 1 + age_centered + I(age_centered^2),
                    data = mdata,
                    family = lognormal(),
                    warmup = 2000, 
                    iter   = 6000, 
                    chains = 4, 
                    cores  = 4, 
                    threads = threading(8), 
                    control = list(adapt_delta = 0.95), 
                    backend = "cmdstanr") %>% add_criterion(c("loo","waic"))

mquadratic

pp_check(mquadratic)

conditional_effects(mquadratic, ndraws = 200, spaghetti = TRUE) %>% 
  plot(points = T)

write.csv(mquadratic, "../saves/lifespan_mquadratic.csv", quote = F, row.names = F)
saveRDS(mquadratic, "../saves/lifespan_mquadratic.rds")

## cubic age
mcubic <- brm(mean_imprecision ~ 1 + age_centered + I(age_centered^2) + I(age_centered^3),
                    data = mdata,
                    family = lognormal(),
                    warmup = 2000, 
                    iter   = 6000, 
                    chains = 4, 
                    cores  = 4, 
                    threads = threading(8), 
                    control = list(adapt_delta = 0.95), 
                    backend = "cmdstanr") %>% add_criterion(c("loo","waic"))

mcubic

pp_check(mcubic)

plot(mcubic)

conditional_effects(mcubic, ndraws = 200, spaghetti = TRUE) %>% 
  plot(points = T)

write.csv(mcubic, "../saves/lifespan_mcubic.csv", quote = F, row.names = F)
saveRDS(mcubic, "../saves/lifespan_mcubic.rds")

## gaussian process model
mgaussianprocess <- brm(mean_imprecision ~ gp(age_centered, k=50, c=5/4, scale=TRUE),
                    data = mdata,
                    family = lognormal(),
                    warmup = 5000, 
                    iter   = 20000, 
                    chains = 4, 
                    cores  = 4, 
                    threads = threading(8), 
                    control = list(adapt_delta = 0.95), 
                    backend = "cmdstanr")

mgaussianprocess <- mgaussianprocess %>% add_criterion(c("loo","waic"))

pp_check(mgaussianprocess)

conditional_effects(mgaussianprocess, ndraws = 200, spaghetti = TRUE) %>% 
  plot(points = T)

write.csv(mgaussianprocess, "../saves/lifespan_mgaussianprocess.csv", quote = F, row.names = F)
saveRDS(mgaussianprocess, "../saves/lifespan_mgaussianprocess.rds")
```

```{r modelcomparison}
# mlinear <- readRDS("../saves/lifespan_mlinear.rds")
# mquadratic <- readRDS("../saves/lifespan_mquadratic.rds")
# mcubic <- readRDS("../saves/lifespan_mcubic.rds")
# mgaussianprocess <- readRDS("../saves/lifespan_mgaussianprocess.rds")

mcomparison <- loo_compare(mlinear, mquadratic, mcubic, mgaussianprocess, criterion = "waic") %>% 
  as_tibble(rownames = "model") %>%
  left_join(as_tibble(model_weights(mlinear, mquadratic, mcubic, mgaussianprocess, weights = "waic"), rownames = "model")) %>% 
  rename(modelweight = value)

mcomparison %>% 
  select(-c(se_elpd_waic, p_waic, se_p_waic)) %>% 
  mutate(across(where(is.numeric), round, 4))

write.csv(mcomparison, "../saves/lifespan_mcomparison.csv", quote = F, row.names = F)
saveRDS(mcomparison, "../saves/lifespan_mcomparison.rds")
```

```{r draws_winningmodel}
# mgaussianprocess <- readRDS("../saves/lifespan_mgaussianprocess.rds")

# mdraws <- mdata %>%
#   add_predicted_draws(mgaussianprocess) %>% 
#   ungroup() %>% 
#   select(-c(.row, .chain, .iteration)) %>%
#   rename(draw = .draw, predicted_value = .prediction)
# 
# write.csv(mdraws, "../saves/lifespan_mdraws.csv", quote = F, row.names = F)
# saveRDS(mdraws, "../saves/lifespan_mdraws.rds")

nd <- mdata %>% 
  select(-c(mean_imprecision))

pred <- fitted(mgaussianprocess, newdata = nd, re_formula = NA) %>% as_tibble() %>% bind_cols(nd) 
```

# CHANGE POINT ANALYSIS

```{r changepoints}
x <- mdata %>% arrange(age) %>% pull(age)
y <- mdata %>% arrange(age) %>% pull(mean_imprecision)

## irregular time intervals
ibeast <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0), # for constant
            tcp.minmax = c(0, 5), # we assume 3, allow some room for exploration
            tseg.min = 10, 
            mcmc.chains = 4, mcmc.burnin = 1000, mcmc.samples = 10000, 
            ci = TRUE)

print(ibeast)
plot(ibeast)

ibeast$trend$cp
```

# PLOT

```{r plot}
ggplot() +

  geom_jitter(
    data = mdata, 
    aes(x = age, y = mean_imprecision), 
    col = "darkgrey", alpha = 0.75
  ) +

  geom_smooth(
    data = pred, 
    aes(x = age, y = Estimate, ymin = Q2.5, ymax = Q97.5), 
    stat = "identity", size = 1, alpha = 0.15, fill = "#006C66", col = "#006C66",
  ) +

  # TODO: add lines where changepoints
  scale_x_continuous(limits = c(0, 80), breaks = seq(0, 80, 5)) + 
  scale_y_continuous(limits = c(0, 4), breaks = seq(0, 4, 0.5)) +

  labs(x = "Age in years", y = "Average imprecision in target width")
```