### 2023-07-12 stats/supplements_changepoint_parameters.Rmd
update lifespan analysis, move around cogn model

---
title: "Change point analysis"
author: "Julia Prein"
date: "03/21/2023"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse) # data handling etc.
library(tidybayes) # for stat_halfeye, add_predictive_draws
library(tidyboot) # for bootstrapped CIs
library(brms) # modeling
library(cmdstanr) # modeling
library(Rbeast) # Bayesian change point analysis
check_cmdstan_toolchain(fix = TRUE, quiet = TRUE)

options(scipen = 999)
theme_set(theme_classic())

# Seed for random number generation
set.seed(42)
```

### Data preparation

```{r load_data}
lifespan_data <- readRDS("../data/tango-lifespan.rds")
```

```{r prepare_modeldata}
mdata <- lifespan_data %>% 
  group_by(subj_id, age, age_centered) %>% 
  summarize(mean_imprecision = mean(click_dist_balloons, na.rm = TRUE))

x <- mdata %>% arrange(age) %>% pull(age)
y <- mdata %>% arrange(age) %>% pull(mean_imprecision)
```

### BEASTs

```{r run_beasts, echo=T, message=FALSE, warning=FALSE, results=FALSE}
b1 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0), 
            tcp.minmax = c(0, 5),
            tseg.min = 10, 
            mcmc.chains = 4, mcmc.burnin = 1000, mcmc.samples = 10000, 
            ci = TRUE)

b2 <- beast.irreg(y = y, time = x, season = "none", 
            mcmc.chains = 4, mcmc.burnin = 1000, mcmc.samples = 10000, 
            ci = TRUE)

b3 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0),
            mcmc.chains = 4, mcmc.burnin = 1000, mcmc.samples = 10000, 
            ci = TRUE)

b4 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0),
            tcp.minmax = c(0, 10), 
            mcmc.chains = 4, mcmc.burnin = 1000, mcmc.samples = 10000, 
            ci = TRUE)

b5 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0),
            tcp.minmax = c(0, 10), 
            tseg.min = 50, 
            mcmc.chains = 4, mcmc.burnin = 1000, mcmc.samples = 10000, 
            ci = TRUE)

b6 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0),
            tcp.minmax = c(0, 10), 
            tseg.min = 100, 
            mcmc.chains = 4, mcmc.burnin = 1000, mcmc.samples = 10000, 
            ci = TRUE)

b7 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0),
            tcp.minmax = c(0, 3), 
            mcmc.chains = 4, mcmc.burnin = 1000, mcmc.samples = 10000, 
            ci = TRUE)

b8 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0),
            tcp.minmax = c(0, 20), 
            mcmc.chains = 4, mcmc.burnin = 1000, mcmc.samples = 10000, 
            ci = TRUE)

b9 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0),
            tcp.minmax = c(0, 100), 
            mcmc.chains = 4, mcmc.burnin = 1000, mcmc.samples = 10000, 
            ci = TRUE)

b10 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0),
            tcp.minmax = c(0, 100), 
            tseg.min = 100, 
            mcmc.chains = 4, mcmc.burnin = 1000, mcmc.samples = 10000, 
            ci = TRUE)
```

```{r fct_printbeasts}
showBeast <- function(beast) {

  print(beast)
  plot(beast)
  print("...continuing to next beast")
}
```

```{r actually_printbeasts}
beasts <- list(b1, b2, b3, b4, b5, b6, b7, b8, b9, b10)

counter = 1

for (i in beasts) {
  print("")
  print("This is the BEAST...")
  print(counter)
  print("")

  showBeast(i)
  counter = counter + 1 
}
```

### Searching for reasonable minimum nr between change points

```{r}
mdata <- mdata %>% 
  ungroup() %>% 
  arrange(age) %>% 
  mutate(
    age_diff = lead(age) - age, 
    age_years = floor(age), 
    three_year_bins = case_match(age_years %% 3, 
                                 0 ~ age_years,
                                 1 ~ age_years-1,
                                 2 ~ age_years-2),
  ) %>% 
  select(-age_years)

mean_age_diff <- mean(mdata$age_diff, na.rm = T) %>% round(2)

mean_n_per_3y <- mdata %>% 
  group_by(three_year_bins) %>% 
  summarize(n_per_bin = n()) %>% 
  summarize(mean_n_per_3y = mean(n_per_bin) %>% round(2)) %>% 
  pull(mean_n_per_3y)

mean_age_diff * mean_n_per_3y * 3

3/mean_age_diff
```

```{r final_beast, message=FALSE, warning=FALSE}
bfinal <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0), 
            tcp.minmax = c(0, 5),
            tseg.min = mean_age_diff * mean_n_per_3y * 3, 
            mcmc.chains = 4, mcmc.burnin = 1000, mcmc.samples = 10000, 
            ci = TRUE)
```

```{r final_beast}
showBeast(bfinal)

# pull out all relevant change points
cps <- tibble(
          cp = bfinal$trend$cp[1:bfinal$trend$ncp_median], 
          lowerCrI = bfinal$trend$cpCI[1:bfinal$trend$ncp_median,1], 
          upperCrI = bfinal$trend$cpCI[1:bfinal$trend$ncp_median,2]
       )
```