### 2023-07-19 stats/lifespan.Rmd
adjust lifespan plot (blue spaghetti)

---
title: "Lifespan analysis"
author: "Julia Prein"
date: "03/21/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse) # data handling etc.
library(tidybayes) # for stat_halfeye, add_predictive_draws
library(tidyboot) # for bootstrapped CIs
library(brms) # modeling
library(cmdstanr) # modeling
library(Rbeast) # Bayesian change point analysis
check_cmdstan_toolchain(fix = TRUE, quiet = TRUE)

options(scipen = 999)
theme_set(theme_classic())

# Seed for random number generation
set.seed(42)
```

```{r load_data}
lifespan_data <- readRDS("../data/tango-lifespan.rds")
```

```{r prepare_modeldata}
mdata <- lifespan_data %>% 
  group_by(subj_id, age, age_centered) %>% 
  summarize(mean_imprecision = mean(click_dist_balloons, na.rm = TRUE))
```

```{r sample}
lifespan_sample <- lifespan_data %>% 
  group_by(age_group) %>%
  mutate(female = ifelse(gender == "f", 1, 0)) %>% 
  summarise(
    n_total = n_distinct(subj_id), 
    age_mean = mean(age, na.rm = T) %>% round(2),
    age_sd = sd(age, na.rm = T) %>% round(2),
    age_min = min(age, na.rm = T),
    age_max = max(age, na.rm = T),
    n_females = ceiling(sum(female/15)),
  )

lifespan_sample

write.csv(lifespan_sample, "../saves/lifespan_sample.csv", quote = F, row.names = F)
saveRDS(lifespan_sample, "../saves/lifespan_sample.rds")
```

```{r variation}
lifespan_data %>% 
  group_by(age_group) %>%
  summarise(
    imprecision_mean = mean(click_dist_balloons, na.rm = T) %>% round(2),
    imprecision_sd = sd(click_dist_balloons, na.rm = T) %>% round(2),
    imprecision_min = min(click_dist_balloons, na.rm = T),
    imprecision_max = max(click_dist_balloons, na.rm = T),
  )
```

# MODEL COMPARISON FOR DEVELOPMENTAL TRAJECTORY

```{r models, message=FALSE, warning=FALSE}
# linear age
mlinear <- brm(mean_imprecision ~ age_centered,
                    data = mdata,
                    family = lognormal(),
                    warmup = 5000, 
                    iter   = 20000, 
                    chains = 4, 
                    cores  = 4, 
                    threads = threading(8), 
                    control = list(adapt_delta = 0.95), 
                    backend = "cmdstanr") %>% add_criterion(c("loo","waic"))

mlinear

pp_check(mlinear)

write.csv(mlinear, "../saves/lifespan_mlinear.csv", quote = F, row.names = F)
saveRDS(mlinear, "../saves/lifespan_mlinear.rds")

## quadratic age
mquadratic <- brm(mean_imprecision ~ 1 + age_centered + I(age_centered^2),
                    data = mdata,
                    family = lognormal(),
                    warmup = 5000, 
                    iter   = 20000, 
                    chains = 4, 
                    cores  = 4, 
                    threads = threading(8), 
                    control = list(adapt_delta = 0.95), 
                    backend = "cmdstanr") %>% add_criterion(c("loo","waic"))

mquadratic

pp_check(mquadratic)

conditional_effects(mquadratic, ndraws = 200, spaghetti = TRUE) %>% 
  plot(points = T)

write.csv(mquadratic, "../saves/lifespan_mquadratic.csv", quote = F, row.names = F)
saveRDS(mquadratic, "../saves/lifespan_mquadratic.rds")

## cubic age
mcubic <- brm(mean_imprecision ~ 1 + age_centered + I(age_centered^2) + I(age_centered^3),
                    data = mdata,
                    family = lognormal(),
                    warmup = 5000, 
                    iter   = 20000, 
                    chains = 4, 
                    cores  = 4, 
                    threads = threading(8), 
                    control = list(adapt_delta = 0.95), 
                    backend = "cmdstanr") %>% add_criterion(c("loo","waic"))

mcubic

pp_check(mcubic)

plot(mcubic)

conditional_effects(mcubic, ndraws = 200, spaghetti = TRUE) %>% 
  plot(points = T, yaxp = c(0,4,1))

write.csv(mcubic, "../saves/lifespan_mcubic.csv", quote = F, row.names = F)
saveRDS(mcubic, "../saves/lifespan_mcubic.rds")

## gaussian process model
mgaussianprocess <- brm(mean_imprecision ~ gp(age_centered, k=50, c=5/4, scale=TRUE),
                    data = mdata,
                    family = lognormal(),
                    warmup = 5000, 
                    iter   = 20000, 
                    chains = 4, 
                    cores  = 4, 
                    threads = threading(8), 
                    control = list(adapt_delta = 0.95), 
                    backend = "cmdstanr")

mgaussianprocess <- mgaussianprocess %>% add_criterion(c("loo","waic"))

pp_check(mgaussianprocess)

conditional_effects(mgaussianprocess, ndraws = 200, spaghetti = TRUE) %>% 
  plot(points = T)

write.csv(mgaussianprocess, "../saves/lifespan_mgaussianprocess.csv", quote = F, row.names = F)
saveRDS(mgaussianprocess, "../saves/lifespan_mgaussianprocess.rds")
```

```{r modelcomparison}
# mlinear <- readRDS("../saves/lifespan_mlinear.rds")
# mquadratic <- readRDS("../saves/lifespan_mquadratic.rds")
# mcubic <- readRDS("../saves/lifespan_mcubic.rds")
# mgaussianprocess <- readRDS("../saves/lifespan_mgaussianprocess.rds")

mcomparison <- loo_compare(mlinear, mquadratic, mcubic, mgaussianprocess, criterion = "waic") %>% 
  as_tibble(rownames = "model") %>%
  left_join(as_tibble(model_weights(mlinear, mquadratic, mcubic, mgaussianprocess, weights = "waic"), rownames = "model")) %>% 
  rename(modelweight = value)

mcomparison %>% 
  select(-c(se_elpd_waic, p_waic, se_p_waic)) %>% 
  mutate(across(where(is.numeric), round, 4))

write.csv(mcomparison, "../saves/lifespan_mcomparison.csv", quote = F, row.names = F)
saveRDS(mcomparison, "../saves/lifespan_mcomparison.rds")
```

```{r draws_winningmodel}
# mgaussianprocess <- readRDS("../saves/lifespan_mgaussianprocess.rds")

nd <- mdata %>% 
  select(-c(mean_imprecision))

pred <- fitted(mgaussianprocess, newdata = nd, re_formula = NA) %>% 
  as_tibble() %>% 
  bind_cols(nd)

pred <- pred %>% clean_names()

write.csv(pred, "../saves/lifespan_mgpdraws.csv", quote = F, row.names = F)
saveRDS(pred, "../saves/lifespan_mgpdraws.rds")
```

# CHANGE POINT ANALYSIS

```{r changepoints}
x <- mdata %>% arrange(age) %>% pull(age)
y <- mdata %>% arrange(age) %>% pull(mean_imprecision)

## irregular time intervals
ibeast <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0), # for constant
            tcp.minmax = c(0, 5), # we assume 3, allow some room for exploration
            tseg.min = 10, 
            mcmc.chains = 4, mcmc.burnin = 1000, mcmc.samples = 10000, 
            ci = TRUE)

print(ibeast)
plot(ibeast)

# how many changepoints?
ibeast$trend$ncp_median

# location of changepoints
ibeast$trend$cp

# pull out all relevant change points
cps <- tibble(
          cp = ibeast$trend$cp[1:ibeast$trend$ncp_median], 
          lowerCrI = ibeast$trend$cpCI[1:ibeast$trend$ncp_median,1], 
          upperCrI = ibeast$trend$cpCI[1:ibeast$trend$ncp_median,2]
       )

saveRDS(ibeast, "../saves/lifespan_ibeast.rds")

write.csv(cps, "../saves/lifespan_ibeast_cps.csv", quote = F, row.names = F)
saveRDS(cps, "../saves/lifespan_ibeast_cps.rds")
```

# PLOT
```{r plot}
# mgaussianprocess <- readRDS("../saves/lifespan_mgaussianprocess.rds")
# cps <- readRDS("../saves/lifespan_ibeast_cps.rds")

# https://cran.r-project.org/web/packages/tidybayes/vignettes/tidy-brms.html
mdata %>%
  add_epred_draws(mgaussianprocess, ndraws = 100) %>%

  ggplot(aes(x = age, y = mean_imprecision)) +

  geom_jitter(data = mdata, aes(x = age, y = mean_imprecision), alpha = 0.5, color = "darkgrey") +

  geom_line(aes(y = .epred, group=.draw), alpha = .075, color = "blue") +

  geom_smooth(data = pred, aes(x = age, y = estimate), stat = "identity", col = "black", size = 0.75) +

  geom_vline(xintercept = cps$cp, linetype = "dashed", alpha = 0.9, size = 0.75) +

  scale_x_continuous(limits = c(0, 80), breaks = seq(0, 80, 5)) +
  scale_y_continuous(limits = c(0, 4), breaks = seq(0, 4, 0.5)) +

  # ## to zoom in
  # coord_cartesian(xlim = c(3, 10)) +
  # scale_x_continuous(limits = c(0, 80), breaks = seq(0, 80, 1)) +

  labs(x = "Age in years", y = "Average imprecision in target width")

# ggsave("../figures/lifespan.png", width = 7, height = 4)
# ggsave("../figures/lifespan.pdf", width = 7, height = 4)
```