### 2023-08-02 stats/magnet.Rmd

---
title: "Magnet analysis"
author: "Julia Prein"
date: "06/07/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse) # data handling etc.
library(psych) # corr.test
library(brms) # modeling
library(tidybayes) # modeling
library(ggpubr) # stat_cor 
library(corrplot) # corr plot
library(apaTables) # as name says
library(cocor) # compare correlations
library(magrittr) # extract list elements
library(cowplot) # ggdraw, draw_image, plot_grid

options(scipen = 999)
theme_set(theme_classic())

# Seed for random number generation
set.seed(42)
```

```{r load_data}
magnet_data <- readRDS("../data/magnet-data.rds")
```

```{r sample}
magnet_sample <- magnet_data %>% 
  mutate(female = ifelse(gender == "f", 1, 0)) %>% 
  summarise(
    n_total = n_distinct(subj_id), 
    age_mean = mean(age, na.rm = T) %>% round(2),
    age_sd = sd(age, na.rm = T) %>% round(2),
    age_min = min(age, na.rm = T),
    age_max = max(age, na.rm = T),
    n_females = sum(female, na.rm = T),
  )

magnet_sample

write.csv(magnet_sample, "../saves/magnet_sample.csv", quote = F, row.names = F)
saveRDS(magnet_sample, "../saves/magnet_sample.rds")
```

```{r scatterplots}
# doesnt make a difference whether we exclude outlier / "imprecise" kids
magnet_data %>% 
  ggplot(., aes(x = magnet_mean, y = tango_mean)) + 
  geom_abline(intercept = 0, slope = 1, col = "darkgrey", linetype = "longdash") +
  geom_smooth(method = "lm", size = 1, alpha = 0.25, col = "darkgrey") +
  geom_point(color = "#005555") +
  stat_cor(method = "pearson", label.x = 5, label.y = 6, aes(label = paste(..r.label..)), size = 5, r.accuracy = 0.01, cor.coef.name = "r") + 
  scale_x_continuous(limits = c(0, 6), breaks = seq(0, 6, 0.5)) +
  scale_y_continuous(limits = c(0, 6), breaks = seq(0, 6, 0.5)) +
  labs(x = "Magnet mean", y = "Tango mean")

magnet_data %>% 
  ggplot(., aes(x = magnet_mean, y = tango_mean)) + 
  geom_abline(intercept = 0, slope = 1, col = "darkgrey", linetype = "longdash") +
  geom_smooth(method = "lm", size = 1, alpha = 0.25, col = "darkgrey") +
  geom_point(color = "#005555") +
  stat_cor(method = "pearson", label.x = 3.5, label.y = 3.25, aes(label = paste(..r.label..)), size = 5, r.accuracy = 0.01, cor.coef.name = "r") + 
  scale_x_continuous(limits = c(0, 3.75), breaks = seq(0, 4, 0.5)) +
  scale_y_continuous(limits = c(0, 3.75), breaks = seq(0, 4, 0.5)) +
  labs(x = "Magnet mean", y = "Tango mean")
```

```{r corrplot}
magnet_data %>% 
  select(tango_mean, magnet_mean, tom_aggregate, perspective_aggregate) %>% 
  rename("Tango mean" = tango_mean, "Magnet mean" = magnet_mean, "ToM score" = tom_aggregate, "Perspective-taking score" = perspective_aggregate) %>% 
  cor() %>% 
  corrplot(diag = FALSE, type = 'lower', method = 'color', order = 'FPC', col = COL2('BrBG'), tl.col = 'black', cl.ratio = 0.2, tl.srt = 0, addCoef.col ='black', number.cex = 0.8)
```

```{r cor_matrix}
tango_magnet <- magnet_data %>% select(tango_mean, magnet_mean) %>% corr.test(method = "pearson")
tango_tom <- magnet_data %>% select(tango_mean, tom_aggregate, perspective_aggregate) %>% corr.test(method = "spearman")
magnet_tom <- magnet_data %>% select(magnet_mean, tom_aggregate, perspective_aggregate) %>% corr.test(method = "spearman")

cor_matrix <- rbind(tango_magnet$ci, tango_tom$ci, magnet_tom$ci) %>% 
  mutate(Correlation = rownames(.)) %>%
  as.tibble()

cor_matrix <- cor_matrix %>%
  # remove tom - perspective correlations
  filter(Correlation != "tm_gg-prsp_1") %>%
  mutate(
    Correlation = recode(Correlation,
                  `tng_m-mgnt_` = "Tango x Magnet", 
                  `tng_m-tm_gg` = "Tango x ToM",
                  `tng_m-prsp_` = "Tango x Perspective-taking",
                  `mgnt_-tm_gg` = "Magnet x ToM",
                  `mgnt_-prsp_` = "Magnet x Perspective-taking", 
                  `tm_gg-prsp_` = "ToM x Perspective-taking")
  ) %>% 
  mutate(across(where(is.numeric), round, 2)) %>% 
  select(Correlation, r, lower, upper, p)

# apa table function only supports pearson corr so far
# magnet_data %>% select(tango_mean, magnet_mean, tom_aggregate, perspective_aggregate) %>% apa.cor.table(filename = "../saves/magnet_corrtable.doc", table.number=1)

cor_matrix

write.csv(cor_matrix, "../saves/magnet_cormatrix.csv", quote = F, row.names = F)
saveRDS(cor_matrix, "../saves/magnet_cormatrix.rds")
```

```{r plot_cor_ci}
cor_ci <- cor_matrix %>% 
  filter(Correlation != "ToM x Perspective-taking") %>% 
  mutate(Correlation = fct_relevel(Correlation, "Magnet x Perspective-taking", "Magnet x ToM", "Tango x Perspective-taking", "Tango x ToM", "Tango x Magnet")) %>%
  ggplot(., aes(x=Correlation, y=r, ymin=lower, ymax=upper)) +
  geom_pointrange(color = "black", size = 0.8, linewidth = 1) +
  scale_y_continuous(limits = c(-1, 1), breaks = seq(-1, 1, 0.25)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab("") +
  coord_flip() +
  theme(axis.title.x = element_text(face="italic"))

cor_ci

saveRDS(cor_ci, file = "../saves/magnet_cor_ci.rds")
# ggsave("../figures/magnet_cor_ci.png", width = 5, height = 3)
```

```{r check_tango_distribution}
ggplot(magnet_data, aes(x = tango_mean)) + 
  geom_density() + 
  stat_function(fun = dnorm, 
                args = with(magnet_data, c(mean = mean(tango_mean), sd = sd(tango_mean))), 
                color = "red", linetype = "dotted") +

  stat_function(fun = dlnorm, 
                args = with(magnet_data, c(meanlog = mean(log(tango_mean)), sdlog = sd(log(tango_mean)))),
                color = "blue", linetype = "dotted")
```

```{r prepare_modeldata}
mdata <- magnet_data %>% 
  mutate(
    magnet_scaled = scale(magnet_mean), 
    tom_scaled = scale(tom_aggregate), 
    perspective_scaled = scale(perspective_aggregate)
  ) %>% 
  select(subj_id, tango_mean, age_centered, magnet_scaled, tom_scaled, perspective_scaled)
```

```{r models, message=FALSE, warning=FALSE}
mage <- brm(tango_mean ~ age_centered,
                    data = mdata,
                    family = lognormal(),
                    warmup = 1000, 
                    iter   = 3000, 
                    chains = 4, 
                    inits  = "random",
                    cores  = 4) %>%
  add_criterion(c("loo","waic"))

mmagnet <- update(mage, formula. = ~ . + magnet_scaled, newdata = mdata) %>% 
  add_criterion(c("loo","waic"))

mtom <- update(mage, formula. = ~ . + tom_scaled, newdata = mdata) %>% 
  add_criterion(c("loo","waic"))

mperspective <- update(mage, formula. = ~ . + perspective_scaled, newdata = mdata) %>% 
  add_criterion(c("loo","waic"))

mmagnet_tom <- update(mage, formula. = ~ . + magnet_scaled + tom_scaled, newdata = mdata) %>% 
  add_criterion(c("loo","waic"))

mmagnet_perspective <- update(mage, formula. = ~ . + magnet_scaled + perspective_scaled, newdata = mdata) %>%
  add_criterion(c("loo","waic"))
```

```{r modelcomparison}
mcomparison <- loo_compare(mage, mmagnet, mtom, mperspective, mmagnet_tom, mmagnet_perspective, criterion = "waic") %>% as_tibble(rownames = "model") %>%
  left_join(as_tibble(model_weights(mage, mmagnet, mtom, mperspective, mmagnet_tom, mmagnet_perspective, weights = "waic"), rownames = "model")) %>% 
  rename(modelweight = value)

mcomparison

model_results <- mcomparison %>%
  select(model, waic, se_waic, modelweight, elpd_diff, se_diff) %>%
  rename(
    Model = model, 
    WAIC = waic, 
    SE_WAIC = se_waic,
    Weight = modelweight, 
    ELPD_DIFF = elpd_diff, 
    SE_ELPD_DIFF = se_diff,
  ) %>% 
  mutate(
    Model = recode(Model, 
                   "mage" = "Null model with Age (scaled)", 
                   "mmagnet" = "Magnet mean (scaled)", 
                   "mtom" = "ToM aggregate (scaled)", 
                   "mperspective" = "Perspective-taking aggregate (scaled)", 
                   "mmagnet_tom" = "Magnet mean (scaled) + ToM aggregate (scaled)", 
                   "mmagnet_perspective" = "Magnet mean (scaled) + Perspective-taking aggregate (scaled)", 
                   )
  )

write.csv(model_results, "../saves/magnet_modelresults.csv", quote = F, row.names = F)
saveRDS(model_results, "../saves/magnet_modelresults.rds")

write.csv(mmagnet_perspective, "../saves/magnet_winningmodel.csv", quote = F, row.names = F)
saveRDS(mmagnet_perspective, "../saves/magnet_winningmodel.rds")
```

```{r magnet_draws}
magnet_draws <-mmagnet_perspective %>% 
  gather_draws(b_age_centered, b_magnet_scaled, b_perspective_scaled) %>% 
  mutate(
    .variable = factor(.variable, 
                       levels = c("b_age_centered", "b_perspective_scaled", "b_magnet_scaled"), 
                       labels = c("Age (scaled)", "Perspective-taking (scaled)", "Magnet mean (scaled)")), 
  ) %>% 
  rename(estimate = .value, predictor = .variable)

write.csv(magnet_draws, "../saves/magnet_draws.csv", quote = F, row.names = F)
saveRDS(magnet_draws, "../saves/magnet_draws.rds")
```

```{r modelplot}
magnet_draws <- readRDS("../saves/magnet_draws.rds")

modelplot <- magnet_draws %>% 
  mutate(
    includes_zero = ifelse(predictor == "Age (scaled)", TRUE, FALSE)
  ) %>%
  ggplot(., aes(x = estimate, y = predictor, fill = includes_zero)) +
 tidybayes::stat_halfeye(.width = c(.95, .8), scale = 1, slab_color = "darkgrey", slab_alpha = 0.9) +
    labs(x = "Posterior estimate TANGO mean", y = "") +
  guides(fill = "none")+
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_fill_manual(values = c("darkgrey", "#EBECEC")) +
  coord_cartesian(expand = TRUE, xlim = c(-0.6, 0.4), ylim = c(1.5, 3.5)) +
  theme_classic()

modelplot

saveRDS(modelplot, file = "../saves/magnet_modelplot.rds")
# ggsave("../figures/modelplot.png", width = 10, height = 6, scale = 0.7, bg = "white")
```

```{r williamstest}
williams_tom <- cocor.dep.groups.overlap(
                         r.jk = cor_matrix$r[cor_matrix$Correlation == "Tango x ToM"], 
                         r.jh = cor_matrix$r[cor_matrix$Correlation == "Magnet x ToM"], 
                         r.kh = cor_matrix$r[cor_matrix$Correlation == "Tango x Magnet"], 
                         n = nrow(magnet_data), 
                         alternative="two.sided", alpha=0.05, conf.level=0.95, null.value=0, 
                         return.htest = T) %>% 
                       extract2("williams1959")

saveRDS(williams_tom, "../saves/magnet_williams_tom.rds")

williams_perspective <- cocor.dep.groups.overlap(
                         r.jk = cor_matrix$r[cor_matrix$Correlation == "Tango x Perspective-taking"], 
                         r.jh = cor_matrix$r[cor_matrix$Correlation == "Magnet x Perspective-taking"], 
                         r.kh = cor_matrix$r[cor_matrix$Correlation == "Tango x Magnet"], 
                         n = nrow(magnet_data), 
                         alternative="two.sided", alpha=0.05, conf.level=0.95, null.value=0, 
                         return.htest = T) %>% 
                       extract2("williams1959")

saveRDS(williams_perspective, "../saves/magnet_williams_perspective.rds")
```

```{r arrange_plot}
procedure_png <- ggdraw() + draw_image("../figures/procedure.png")
cor_ci <- readRDS(file = "../saves/magnet_cor_ci.rds")
modelplot <- readRDS(file = "../saves/magnet_modelplot.rds")

magnet_plot <- ggarrange(NULL, 
          procedure_png, 
          cor_ci, 
          modelplot, 
          NULL, 
          ncol = 5, 
          labels = c("A", "", "B", "C", ""),
          widths = c(1, 16, 18, 13, 1.5)
          )

saveRDS(magnet_plot, file = "../saves/magnet_plot.rds")
ggsave("../figures/magnet_plot.png", width = 10, height = 2, scale = 1.4, bg = "white")
ggsave("../figures/magnet_plot.pdf", width = 10, height = 2, scale = 1.4, bg = "white")

# alternative positioning: procedure on top, cor left below, model right below
# ggarrange(procedure_png, 
#           NULL, 
#           
#           ggarrange(
#             cor_ci, 
#             modelplot, 
#             NULL, 
#             ncol = 3, 
#             labels = c("B", "C", ""), 
#             widths = c(12, 10, 2)
#           ), 
#           
#           nrow = 3,
#           heights = c(11, 1, 9), 
#           labels = c("A", ""), 
#           align = "h"
#           )
# 
# # saveRDS(magnet_plot, file = "../saves/magnet_plot.rds")
# ggsave("../figures/magnet_plot2.png", width = 8, height = 5, scale = 1.2, bg = "white")
```