### 2023-10-04 stats/gazemodel.Rmd

---
title: "Gaze model analysis"
author: "Julia Prein"
date: "06/07/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse) # data handling etc.
library(tidyboot) # for bootstrap CIs
library(coda) # HPDinterval
library(ggpubr) # stat_cor
library(ggdist) # stat_pointinterval
library(matrixStats) # logSumExp
library(ggridges) # for geom_density_ridges
library(ggthemes) # themes for plotting
library(cowplot) # ggdraw, draw_image, plot_grid
library(psych) #corr.test

options(scipen = 999)
theme_set(theme_classic())

# Seed for random number generation
set.seed(42)
```

```{r functions}
estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}

hdi_upper<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}

hdi_lower<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}
```

```{r load_data}
tango_kids <- readRDS("../data/tango-model-kids.rds")
tango_adults <- readRDS("../data/tango-model-adults.rds") %>% mutate(age_group = 18)
tango <- full_join(tango_kids, tango_adults)

# careful, this takes a while... 
model_kids <- readRDS("../data/gazemodel-kids.rds")
model_kids_dev_prior <- readRDS("../data/gazemodel-kids-dev-prior.rds")

model_adults_ind <- readRDS("../data/gazemodel-adults-ind.rds") %>% 
  rename("mode" = sdTol) %>% 
  mutate(
    age_group = 18, 
    age = 18,
    id = str_replace_all(id, "gafo_prolific_modelvalidation_vp", "gp1-")
  )

model_adults_bins <- readRDS("../data/gazemodel-adults-bins.rds") %>% 
  mutate(
    age_group = 18, 
    # transform into target width units
    data_mean = data_mean / 160, 
    model_mean = model_mean / 160,
  )
```

```{r sample}
gazemodel_sample_kids <- tango_kids %>% 
  group_by(age_group) %>%
  mutate(female = ifelse(gender == "f", 1, 0)) %>% 
  summarise(
    n_total = n_distinct(subj_id), 
    age_mean = mean(age, na.rm = T) %>% round(2),
    age_sd = sd(age, na.rm = T) %>% round(2),
    age_min = min(age, na.rm = T),
    age_max = max(age, na.rm = T),
    n_females = ceiling(sum(female/15)),
  )

gazemodel_sample_kids

gazemodel_sample_adults <- tango_adults %>% 
  mutate(female = ifelse(gender == "f", 1, 0)) %>% 
  summarise(
    n_total = n_distinct(subj_id), 
    age_mean = mean(age, na.rm = T) %>% round(2),
    age_sd = sd(age, na.rm = T) %>% round(2),
    age_min = min(age, na.rm = T),
    age_max = max(age, na.rm = T),
    n_females = ceiling(sum(female/15)),
  ) %>% 
  mutate(age_group = 18)

gazemodel_sample <- gazemodel_sample_kids %>% full_join(gazemodel_sample_adults)

write.csv(gazemodel_sample, "../saves/gazemodel_sample.csv", quote = F, row.names = F)
saveRDS(gazemodel_sample, "../saves/gazemodel_sample.rds")
```

```{r data_gazefunnel}
target_position_imprecision <- tango %>%
  group_by(age_group, target_position) %>%
  tidyboot_mean(column = click_dist_balloons, na.rm = TRUE)

target_position_imprecision_ind <- tango %>%
  group_by(subj_id, age_group, target_position) %>%
  summarize(mean = mean(click_dist_balloons, na.rm = TRUE))

gazefunnel <- ggplot() +
  geom_point(data = target_position_imprecision_ind, aes(x = target_position, y = mean, col = as.factor(age_group)), alpha = .1, position = position_dodge(width = 0.3)) +

  geom_pointrange(data = target_position_imprecision, aes(x = target_position, y = mean, ymin = ci_lower, ymax = ci_upper, col = as.factor(age_group)), pch = 18,  size = 1, position = position_dodge(width = .3)) +

  scale_y_continuous(limits = c(0, 7), breaks = seq(0, 7, 1)) +
  scale_color_manual(values = c("#c6d325", "#00b1ea", "#29485d", "darkgrey"), breaks = c("3", "4", "5", "18"), labels = c("3-year-olds", "4-year-olds", "5-year-olds", "Adults")) +
    labs(x = "Target position (binned)", y = "Average imprecision in target widths", col = "Age in years")

gazefunnel

# saveRDS(gazefunnel, "../saves/gazemodel_funnel.rds")
# ggsave("../figures/gazemodel_funnel.png", width = 6, height = 4, scale = 1.2)
```

# Inference

```{r plot3A}
# developmental trajectory model inference parameter
inf_map_devp <- model_kids_dev_prior %>%
  filter(scope == "global", 
         parameter == "inference", 
         type != "sigma") %>%
  group_by(type) %>%
  summarise(mode = estimate_mode(value)) %>%
  pivot_wider(names_from = "type", values_from = "mode") %>%
  expand_grid(age = seq(0, 3, 0.1)) %>%
  mutate(sd = exp(intercept + slope * age))

inf_samp_devp <- model_kids_dev_prior %>%
  filter(scope == "global", 
         parameter == "inference", 
         type != "sigma") %>%
  pivot_wider(names_from = "type", values_from = "value") %>%
  sample_n(1000) %>%
  expand_grid(age = seq(0, 3, 0.1)) %>%
  mutate(sd = exp(intercept + slope * age))

id_devp <- model_kids_dev_prior %>%
  filter(scope == "id", 
         type == "sd") %>%
  group_by(id) %>%
  summarise(mode = estimate_mode(value),
            lci = hdi_lower(value), 
            uci = hdi_upper(value)) %>%
  left_join(tango_kids %>% 
              select(subj_id, age, age_group) %>% 
              rename(id = subj_id) %>% 
              distinct(id, .keep_all = T)) %>%
  ungroup()

model_adults_sd <- model_adults_ind %>% 
  summarize(
    mean_mode = mean(mode), 
    lci = hdi_lower(mode), 
    uci = hdi_upper(mode)
  )

dev_plot <- ggplot() +

  # children
  geom_point(data = id_devp, aes(x = age, y = mode), alpha = 0.75, color = "darkgrey") +
  geom_line(data = inf_samp_devp, aes(x = age + 3, y = sd, group = paste(chain,iteration)), col = "blue", alpha = .01) +
  geom_line(data = inf_map_devp, aes(x = age + 3, y = sd), col = "black", linewidth = 1.25) +

  # adults 
  geom_jitter(data = model_adults_ind, aes(x = 6.5, y = mode), alpha = 0.5, color = "darkgrey",  width = .1) + 
  geom_pointrange(data = model_adults_sd %>% head(1), aes(x = 6.5, y = mean_mode, ymin = lci, ymax = uci), size = 0.5) +

  scale_x_continuous(limits = c(3, 6.7), breaks = seq(3, 6.5, 0.5), labels = c("3.0", "3.5", "4.0", "4.5", "5.0", "5.5", "6.0", "Adults")) + 
  scale_y_continuous(limits = c(0, 6), breaks = seq(0, 6, 0.25)) +
  coord_cartesian(ylim = c(0, 3)) +
  labs(x = "Age in years", y = expression(paste("Mode model parameter (", italic("\U03C3"[v]), ")")))

saveRDS(dev_plot, file = "../saves/gazemodel_dev_plot.rds")

# ggsave("../figures/gazemodel_dev.png", width = 8, height = 5, scale = 1, bg = "white")
# ggsave("../figures/gazemodel_dev.pdf", width = 8, height = 5, scale = 1, bg = "white")
```

## Correlation model vs data individual level / age

```{r plot3B}
comparison_model_data_devp <- model_kids_dev_prior %>%
  filter(scope == "id", type == "sd") %>%
  group_by(id) %>%
  summarise(model_mode = estimate_mode(value)) %>%
  left_join(
    tango_kids %>% 
      group_by(subj_id, age, age_group) %>%
      summarise(data_mean = mean(click_dist_balloons)) %>%
      rename(id = subj_id)
  )

comparison_model_data_adults <- tango_adults %>% 
  group_by(subj_id) %>% 
  summarize(
    data_mean = mean(click_dist_balloons)
  ) %>% 
  left_join(model_adults_ind %>% select(-lci, -uci) %>% rename("model_mode" = mode, "subj_id" = id))

comparison_model_data <- full_join(comparison_model_data_devp, comparison_model_data_adults) %>% 
  mutate(sample = ifelse(age < 18, "children", "adults"))

model_vs_data_plot <- ggplot(comparison_model_data, aes(x = model_mode, y = data_mean, group = sample)) +
  geom_point(aes(col = age)) +
  stat_cor(data = comparison_model_data %>% filter(sample == "children"), aes(label = paste(..r.label..)), method = "pearson", size = 4, r.accuracy = 0.01, cor.coef.name = "r", label.x = 0, label.y = 4.5) +

  stat_cor(data = comparison_model_data %>% filter(sample == "adults"), aes(label = paste(..r.label..)), method = "pearson", size = 4, r.accuracy = 0.01, cor.coef.name = "r", label.x = 0, label.y = 2.25) +

  geom_smooth(method = "lm", col = "darkgrey", alpha = 0.25) +

  labs(x = expression(paste("Mode model parameter (", italic("\U03C3"[v]), ")")), y = "Mean data imprecision", color = "Age") +

  scale_x_continuous(limits = c(0, NA), n.breaks = 7) +
  scale_y_continuous(limits = c(0, NA), n.breaks = 5) +
  scale_color_viridis_c(direction = -1, alpha = 0.9, limits = c(3, 6), breaks = seq(3, 6, 0.5)) +
  theme_few() +
  theme(
    legend.position = c(0.37, 0.3),
    strip.text = element_text(face = "bold", size = 10), 
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8), 
    legend.box.background = element_rect(), 
    legend.key.height = unit(0.5, "cm"),
  ) +
  facet_wrap(~ factor(sample, levels = c("children", "adults"), labels = c("Child sample", "Adult sample")), scales = "free")

model_vs_data_plot

saveRDS(model_vs_data_plot, file = "../saves/gazemodel_model_vs_data.rds")

# ggsave("../figures/gazemodel_model_vs_data.png", width = 8, height = 5, scale = 1, bg = "white")
# ggsave("../figures/gazemodel_model_vs_data.pdf", width = 8, height = 5, scale = 1, bg = "white")
```

# Inference parameters individuals

```{r id_inf}
# prepare data
id_inf <- model_kids_dev_prior %>%
  filter(scope == "id", parameter == "inference") %>% 
  group_by(id) %>% 
  mutate(mean = mean(value, na.rm = T))

idinf_plot <- id_inf %>% 
  slice_sample(n = 2000) %>% # for faster debugging
  full_join((tango_kids %>% select(subj_id, age, age_group)), by = join_by(id == subj_id)) %>% 
  ggplot(aes(x = value, y = reorder(id, mean), fill = age)) +
  geom_density_ridges(bandwidth = .05, rel_min_height = .01, size = .25, scale = 3, alpha = .5) + 
  scale_fill_viridis_c(direction = -1, alpha = 0.9, limits = c(3, 6), breaks = seq(3, 6, 0.5)) +
  scale_x_continuous(breaks = seq(0, 7, 0.5)) +
  coord_cartesian(xlim = c(0, 7)) +
  theme(axis.ticks.y = element_blank(), axis.text.y = element_blank()) + 
  labs(x = "Inference parameter", y = "Participants (ordered by mean)", fill = "Age")

idinf_plot
saveRDS(idinf_plot, "../saves/supplements_gazemodel_idinf_plot.rds")

ggsave("../figures/supplements_gazemodel_indinference.png", width = 5, height = 5)
ggsave("../figures/supplements_gazemodel_indinference.pdf", width = 5, height = 5)
```

```{r faceted_inf, message=FALSE, warning=FALSE}
faceted_inf <- model_kids_dev_prior %>%
  filter(scope == "id", parameter == "inference") %>%
  mutate(id = vctrs::vec_group_id(id)) %>% 
  ggplot() + 
  geom_density(aes(x = value, fill = factor(chain)), alpha = .5) +
  stat_pointinterval(aes(x = value)) +
  facet_wrap(~ id, nrow = 6, scales = "free") +
  scale_fill_viridis_d() +
  scale_x_continuous(limits = c(0, 3)) + 
  scale_y_continuous(limits = function(x){c(0, max(0.1, x))}) +
  labs(x = "Inference parameter value", y = "", fill = "Model chains") +
  theme_minimal()

faceted_inf
saveRDS(faceted_inf, "../saves/supplements_gazemodel_facetedinference.rds")

ggsave("../figures/supplements_gazemodel_facetedinference.png", width = 10, height = 6, scale = 1.4, bg = "white")
ggsave("../figures/supplements_gazemodel_facetedinference.pdf", width = 10, height = 6, scale = 1.4, bg = "white")
```

# Recover U pattern

```{r get_click_prior}
get_click_prior <- function(sdTol, eyeCenterLY, eyeCenterLX, eyeCenterRY, eyeCenterRX, targetCenterY, targetCenterX, pupilLX, pupilRX, priorInt, priorSlope, age){

  if(eyeCenterLX > pupilLX) {
    tAlphaL = atan(abs(eyeCenterLX-targetCenterX)/(eyeCenterLY-targetCenterY))
  } else {
    tAlphaL = -atan(abs(eyeCenterLX-targetCenterX)/(eyeCenterLY-targetCenterY))
  }

  #noisyAlphaL = rnorm(1,tAlphaL, sdPer)

  if(eyeCenterRX > pupilRX) {
    tAlphaR = atan(abs(eyeCenterRX-targetCenterX)/(eyeCenterRY-targetCenterY))
  } else {
    tAlphaR = -atan(abs(eyeCenterRX-targetCenterX)/(eyeCenterRY-targetCenterY))
  }

  #noisyAlphaR = rnorm(1,tAlphaR, sdPer)

  bRDist = c()

  bLDist = c()

  for (x in 1:1920){

  if(eyeCenterLX > x) {
    betaL = atan(abs(eyeCenterLX-x)/(eyeCenterLY-targetCenterY))
  } else {
    betaL = -atan(abs(eyeCenterLX-x)/(eyeCenterLY-targetCenterY))
  }
    betaLDist = dnorm(betaL,tAlphaL,sdTol, log = T)

  if(eyeCenterRX > x) {
    betaR = atan(abs(eyeCenterRX-x)/(eyeCenterRY-targetCenterY))
  } else {
    betaR = -atan(abs(eyeCenterRX-x)/(eyeCenterRY-targetCenterY))
  }
    betaRDist = dnorm(betaR,tAlphaR,sdTol, log = T)

    #betaDist = betaLDist + betaRDist

    #bDist[x] <- exp(betaDist)
    bRDist[x] <- betaRDist

    bLDist[x] <- betaLDist

  }

  xRDist = exp(bRDist-logSumExp(bRDist))

  xLDist = exp(bLDist-logSumExp(bLDist))

  priorSD = priorInt + priorSlope * age

  xDist = (xRDist * xLDist) * dnorm(seq(1:1920), mean = 960, sd = priorSD)

  dist = xDist / sum(xDist)

  click = sample(1:1920, prob = xDist, size = 1)-1

  return(click)
}
```

```{r plot3C}
# u pattern recovery
ind_sd <- model_kids_dev_prior %>%
  filter(scope == "id", type == "sd") %>%
  group_by(id) %>%
  summarise(sdTol = estimate_mode(value)) %>%
  left_join(tango_kids %>% rename(id = subj_id,
                          eyeCenterLY = eye_center_left_y,
                          eyeCenterLX = eye_center_left_x,
                          eyeCenterRY = eye_center_right_y,
                          eyeCenterRX = eye_center_right_x,
                          pupilLX = pupil_final_left_x,
                          pupilRX = pupil_final_right_x,
                          targetCenterX = target_center_x,
                          targetCenterY = target_center_y) %>%
              mutate(clickDist = abs(click_dist_from_target_center_x)) %>%
              select(id, age, age_group, trial_nr, eyeCenterLX, eyeCenterLY, eyeCenterRX, eyeCenterRY, pupilLX, pupilRX, targetCenterY, targetCenterX, clickDist))

ind_sd_long <- ind_sd %>% slice(rep(1:n(), each = 100))

# saveRDS(ind_sd_long, file = "../saves/gazemodel_ind_sd_long.rds")
# ind_sd_long <- readRDS("../saves/gazemodel_ind_sd_long.rds")

priorMode <- model_kids_dev_prior %>%
  filter(scope == "global",
         parameter == "priorSd",
         type != "sigma") %>%
  group_by(type) %>%
  summarise(
    mode = estimate_mode(value), 
    lci = hdi_lower(value), 
    uci = hdi_upper(value))

saveRDS(priorMode, file = "../saves/gazemodel_prior.rds")

# # careful, this runs a while... (ca. 30-45 min for me)

# gen_model_clicks <- ind_sd_long %>%
#   rowwise() %>%
#   mutate(modelClick = get_click_prior(
#           sdTol = sdTol,
#           eyeCenterLY = eyeCenterLY,
#           eyeCenterLX = eyeCenterLX, 
#           eyeCenterRY = eyeCenterRY,
#           eyeCenterRX = eyeCenterRX,
#           targetCenterX = targetCenterX,
#           pupilLX = pupilLX,
#           pupilRX = pupilRX,
#           targetCenterY = targetCenterY,
#           priorInt = priorMode %>% filter(type == "intercept") %>% pull(mode),
#           priorSlope = priorMode %>% filter(type == "slope") %>% pull(mode),
#           age = age
#           ))
# 
# modelpred_gaze <- gen_model_clicks %>%
#   mutate(
#     modelClickDist = abs(modelClick - targetCenterX),
#     targetBin = cut(targetCenterX, 
#                     breaks = seq(from = 0, to = 1920, by = 192),
#                     labels = as.character(1:10))
#   ) %>%
#   group_by(age_group, targetBin) %>%
#   summarise(
#     data_mean = mean(clickDist) / 160, 
#     model_mean = mean(modelClickDist) / 160,
#   )

# saveRDS(modelpred_gaze, "../saves/gazemodel_modelpred_gaze.rds")
modelpred_gaze <- readRDS("../saves/gazemodel_modelpred_gaze.rds")

u_data <- full_join(modelpred_gaze, model_adults_bins)
saveRDS(u_data, "../saves/gazemodel_u_data.rds")

u_plot <- u_data %>%
  pivot_longer(cols = c(data_mean, model_mean), names_to = "source", values_to = "imprecision") %>%
  mutate(
    age_group = factor(age_group, levels = c(3, 4, 5, 18), labels = c("3-year-olds", "4-year-olds", "5-year-olds", "Adults")), 
  ) %>% 
  ggplot(aes(x = targetBin, y = imprecision, col = source)) +
  geom_point(alpha = 0.8) +
  facet_grid(~ age_group) +
  scale_y_continuous(limits = c(0, 5.5), breaks = seq(0, 5.5, 0.5)) +
  scale_color_manual(values = c("darkgrey", "darkblue"), labels = c("Data", "Model prediction")) +
  labs(x = "Target bin", y = "Imprecision", color = "Source") + 
  theme_few() +
  theme(
    legend.position = c(0.881, 0.7), # c(0,0) bottom left, c(1,1) top-right
    strip.text = element_text(face = "bold", size = 10), 
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8), 
    legend.box.background = element_rect(), 
    legend.key.height = unit(0.4, "cm"),
    legend.key.width = unit(-1.2, "cm"),
    plot.margin = margin(0.2, 0, 0.1, 0.2, "cm") # t r b l
  )

u_plot
saveRDS(u_plot, "../saves/gazemodel_u_plot.rds")
# ggsave("../figures/gazemodel_Uplot.png", width = 8, height = 5, scale = 1, bg = "white")
# ggsave("../figures/gazemodel_Uplot.pdf", width = 8, height = 5, scale = 1, bg = "white")
```

```{r modelpred_gaze_plot}
# correlation model predicted imprecision vs data imprecision across target positions
modelpred_gaze_plot <- modelpred_gaze %>% 
  ggplot(., aes(x = model_mean, y = data_mean)) +
  geom_point(aes(col = targetBin)) +
  stat_cor(method = "pearson", label.x = 0.2, label.y = 4.9, aes(label = paste(..r.label..)), size = 4, r.accuracy = 0.01, cor.coef.name = "r") +
  geom_abline(intercept = 0, slope = 1, linetype = 3) +
  scale_color_brewer(palette="BrBG") +
  labs(x = "Mean predicted imprecision", y = "Mean data imprecision", color = "Target bin") + 
  scale_x_continuous(limits = c(0, 5.5), breaks = seq(0, 5, 1)) + 
  scale_y_continuous(limits = c(0, 5.5), breaks = seq(0, 5, 1))

modelpred_gaze_plot
saveRDS(modelpred_gaze_plot, "../saves/gazemodel_modelpred_gaze_plot.rds")
# ggsave("../figures/gazemodel_modelpred_gaze_plot.png", width = 8, height = 5, scale = 1, bg = "white")
# ggsave("../figures/gazemodel_modelpred_gaze_plot.pdf", width = 8, height = 5, scale = 1, bg = "white")
```

# Alternative models click generation

## Center bias model

```{r get_click_center}
get_click_center <- function(){

  xDist = dnorm(seq(1:1920), mean = 960, sd = 160)

  click = sample(1:1920, prob = xDist, size = 1)-1

  return(click)
}
```

```{r center_model}
gen_model_clicks_center <- ind_sd_long %>%
  rowwise()%>%
  mutate(modelClick = get_click_center())

modelpred_center <- gen_model_clicks_center %>%
  mutate(
    modelClickDist = abs(modelClick - targetCenterX),
    targetBin = cut(targetCenterX, 
                    breaks = seq(from = 0, to = 1920, by = 192),
                    labels = as.character(1:10))) %>%
  group_by(age_group, targetBin) %>%
  summarise(
    data_mean = mean(clickDist) / 160, 
    model_mean = mean(modelClickDist) / 160, 
  )

saveRDS(modelpred_center, file = "../saves/gazemodel_modelpred_center.rds")
modelpred_center <- readRDS(file = "../saves/gazemodel_modelpred_center.rds")

# facetted U plot
modelpred_center %>%
  pivot_longer(cols = c(data_mean, model_mean), names_to = "source", values_to = "imprecision") %>%
  mutate(
    age_group = factor(age_group, levels = c(3, 4, 5), labels = c("3-year-olds", "4-year-olds", "5-year-olds")), 
  ) %>% 
  ggplot(aes(x = targetBin, y = imprecision, col = source)) +
  geom_point(alpha = 0.8) +
  facet_grid(~ age_group) +
  scale_y_continuous(limits = c(0, 5.5), breaks = seq(0, 5.5, 0.5)) +
  scale_color_manual(values = c("darkgrey", "darkblue"), labels = c("Data", "Model prediction")) +
  labs(x = "Target bin", y = "Imprecision", color = "Source") + 
  theme_few() +
  theme(
    legend.position = c(0.84 , 0.8), # c(0,0) bottom left, c(1,1) top-right
    legend.background = element_rect(color = "darkgrey", size = 0.1), 
    legend.text = element_text(size = 8),
    strip.text = element_text(face = "bold", size = 12)
  )

# corr plot
modelpred_center_plot <- modelpred_center %>% 
  ggplot(., aes(x = model_mean, y = data_mean)) +
  geom_point(aes(col = targetBin), size = 0.8) +
  stat_cor(method = "pearson", label.x = 0.2, label.y = 4.9, aes(label = paste(..r.label..)), size = 3, r.accuracy = 0.01, cor.coef.name = "r") +
  geom_abline(intercept = 0, slope = 1, linetype = 3) +
  scale_color_brewer(palette="BrBG") +
  labs(x = "Mean predicted imprecision", y = "Mean data imprecision", color = "Target bin") + 
  scale_x_continuous(limits = c(0, 5.5), breaks = seq(0, 5, 1)) + 
  scale_y_continuous(limits = c(0, 5.5), breaks = seq(0, 5, 1))

modelpred_center_plot
saveRDS(modelpred_center_plot, "../saves/gazemodel_modelpred_center_plot.rds")
# ggsave("../figures/gazemodel_modelpred_center_plot.png", width = 10, height = 5, scale = 1.4, bg = "white")
# ggsave("../figures/gazemodel_modelpred_center_plot.pdf", width = 10, height = 5, scale = 1.4, bg = "white")
```

## Random clicks model

```{r get_click_random}
get_click_random <- function(){

  xDist = dunif(seq(1:1920), min = 1, max = 1920)

  click = sample(1:1920, prob = xDist, size = 1)-1

  return(click)
}
```

```{r random_model}
gen_model_clicks_random <- ind_sd_long %>%
  rowwise() %>%
  mutate(modelClick = get_click_random())

modelpred_random <- gen_model_clicks_random %>%
  mutate(
    modelClickDist = abs(modelClick - targetCenterX),
    targetBin = cut(targetCenterX, 
                    breaks = seq(from = 0, to = 1920, by = 192),
                    labels=as.character(1:10))
  ) %>%
  group_by(age_group, targetBin) %>%
  summarise(
    data_mean = mean(clickDist) / 160, 
    model_mean = mean(modelClickDist) / 160
  )

saveRDS(modelpred_random, file = "../saves/gazemodel_modelpred_random.rds")
modelpred_random <- readRDS(file = "../saves/gazemodel_modelpred_random.rds")

# facetted U plot
modelpred_random %>%
  pivot_longer(cols = c(data_mean, model_mean), names_to = "source", values_to = "imprecision") %>%
  mutate(
    age_group = factor(age_group, levels = c(3, 4, 5), labels = c("3-year-olds", "4-year-olds", "5-year-olds")), 
  ) %>% 
  ggplot(aes(x = targetBin, y = imprecision, col = source)) +
  geom_point(alpha = 0.8) +
  facet_grid(~ age_group) +
  scale_y_continuous(limits = c(0, 5.5), breaks = seq(0, 5.5, 0.5)) +
  scale_color_manual(values = c("darkgrey", "darkblue"), labels = c("Data", "Model prediction")) +
  labs(x = "Target bin", y = "Imprecision", color = "Source") + 
  theme_few() +
  theme(
    legend.position = c(0.84 , 0.8), # c(0,0) bottom left, c(1,1) top-right
    legend.background = element_rect(color = "darkgrey", size = 0.1), 
    legend.text = element_text(size = 8),
    strip.text = element_text(face = "bold", size = 12)
  )

# corr plot
modelpred_random_plot <- modelpred_random %>% 
  ggplot(., aes(x = model_mean, y = data_mean)) +
  geom_point(aes(col = targetBin), size = 0.8) +
  stat_cor(method = "pearson", label.x = 0.2, label.y = 4.9, aes(label = paste(..r.label..)), size = 3, r.accuracy = 0.01, cor.coef.name = "r") +
  geom_abline(intercept = 0, slope = 1, linetype = 3) +
  scale_color_brewer(palette="BrBG") +
  labs(x = "Mean predicted imprecision", y = "Mean data imprecision", color = "Target bin") + 
  scale_x_continuous(limits = c(0, 5.5), breaks = seq(0, 5, 1)) + 
  scale_y_continuous(limits = c(0, 5.5), breaks = seq(0, 5, 1))

modelpred_random_plot
saveRDS(modelpred_random_plot, "../saves/gazemodel_modelpred_random_plot.rds")
# ggsave("../figures/gazemodel_modelpred_random_plot.png", width = 10, height = 5, scale = 1.4, bg = "white")
# ggsave("../figures/gazemodel_modelpred_random_plot.pdf", width = 10, height = 5, scale = 1.4, bg = "white")
```

## Combine plots of predictions from 3 models

```{r plot3D}
modelpred_gaze_plot <- readRDS("../saves/gazemodel_modelpred_gaze_plot.rds")
modelpred_center_plot <- readRDS("../saves/gazemodel_modelpred_center_plot.rds")
modelpred_random_plot <- readRDS("../saves/gazemodel_modelpred_random_plot.rds")

corr_model_pred_plot <- ggarrange(

  modelpred_gaze_plot + 
    theme(legend.position = c(0.85, 0.3), 
          legend.title = element_text(size = 8),
          legend.text = element_text(size = 8), 
          legend.box.background = element_rect(), 
          legend.key.height = unit(0.3, "cm"),
          legend.key.width = unit(-1.5, "cm"),
          plot.margin = margin(0.2, -2, 0.1, 0.2, "cm") # t r b l
         ),

  ggarrange(
    modelpred_center_plot + 
      labs(title = "Center bias model", x = "", y = "") +
      theme(legend.position = "none", 
            axis.text = element_text(size = 6), 
            plot.title = element_text(size = 8), 
            plot.subtitle = element_text(size = 8), 
            plot.margin = margin(0.2, 0.2, 0.2, 0, "cm")),

    modelpred_random_plot + 
      labs(title = "Random guessing model", x = "", y = "") +
      theme(legend.position = "none", 
            axis.text = element_text(size = 6), 
            plot.title = element_text(size = 8), 
            plot.subtitle = element_text(size = 8), 
            plot.margin = margin(0.2, 0.2, 0.2, 0, "cm")), 
    nrow = 2
  ),

  ncol = 2, 
  widths = c(1.5, 1), 
  align = "v"
  )

corr_model_pred_plot
saveRDS(corr_model_pred_plot, file = "../saves/gazemodel_corr_model_pred_plot.rds")
```

# Plot

```{r arrange_plot}
dev_plot <- readRDS(file = "../saves/gazemodel_dev_plot.rds")
model_vs_data_plot <- readRDS(file = "../saves/gazemodel_model_vs_data.rds")
u_plot <- readRDS("../saves/gazemodel_u_plot.rds")
corr_model_pred_plot <- readRDS("../saves/gazemodel_corr_model_pred_plot.rds")

gazemodel_plot <- ggarrange(
  dev_plot, 
  NULL, 
  model_vs_data_plot,

  u_plot, 
  NULL, 
  corr_model_pred_plot,

  ncol = 3, 
  nrow = 2, 
  widths = c(13, 1, 12), 
  labels = c("A", "", "B", "C", "", "D")
)

gazemodel_plot
saveRDS(gazemodel_plot, "../saves/gazemodel_plot.rds")
ggsave("../figures/gazemodel_plot.png", width = 8, height = 5, scale = 1.5, bg = "white")
ggsave("../figures/gazemodel_plot.pdf", width = 8, height = 5, scale = 1.4, bg = "white", device = cairo_pdf)
# add cairo_pdf to enable greek symbols in pdf saving
```

```{r results_table}
# create tibble for reading out stats results in rmd manuscript
corr_gazemodel_param_children <- comparison_model_data_devp %>% select(model_mode, data_mean) %>% corr.test(method = "pearson")

corr_gazemodel_param_adults <- comparison_model_data_adults %>% select(model_mode, data_mean) %>% corr.test(method = "pearson")

# this is children only 
corr_gazemodel_pred <- modelpred_gaze %>% ungroup() %>% select(model_mean, data_mean) %>% corr.test(method = "pearson")
corr_centermodel_pred <- modelpred_center %>% ungroup() %>% select(model_mean, data_mean) %>% corr.test(method = "pearson")
corr_randommodel_pred <- modelpred_random %>% ungroup() %>% select(model_mean, data_mean) %>% corr.test(method = "pearson")

corrs <- rbind(corr_gazemodel_param_children$ci, corr_gazemodel_param_adults$ci, corr_gazemodel_pred$ci, corr_centermodel_pred$ci, corr_randommodel_pred$ci) %>% 
  rownames_to_column(var = "corr") %>% 
  mutate(
    corr = recode(
            corr,
            `mdl_m-dt_mn` = "gazemodel_param_kids", 
            `mdl_m-dt_mn1` = "gazemodel_param_adults", 
            `mdl_m-dt_mn2` = "gazemodel_pred_kids", 
            `mdl_m-dt_mn3` = "centermodel_pred_kids", 
            `mdl_m-dt_mn4` = "randommodel_pred_kids", 
           )
  ) %>% 
  mutate(across(where(is.numeric), round, 2))

corrs
saveRDS(corrs, file = "../saves/gazemodel_correlations.rds")
```

# Model comparison BayesFactor

```{r read_data_mcomparison}
modelcomparison_kids_bf <- readRDS("../data/gazemodel-kids-comparison-bf.rds")
modelcomparison_kids_samples <- readRDS("../data/gazemodel-kids-comparison-samples.rds")
```

```{r mcomparison}
# same as reading in modelcomparison_kids_bf
modelcomparison_kids_samples %>%
  group_by(model) %>%
  summarise(margllh = logSumExp(like)) %>%
  pivot_wider(names_from = model, values_from = margllh) %>%
  summarise(bf_inf_center = inf - center, 
            bf_inf_guess = inf - guess)
```

# Headlight distribution for plots

```{r get_dist}
# get_dist <- function(sdTol, eyeCenterLY, eyeCenterLX, eyeCenterRY, eyeCenterRX, targetCenterY, targetCenterX, pupilLX, pupilRX){
# 
#   if(eyeCenterLX > pupilLX) {
#     tAlphaL = atan(abs(eyeCenterLX-targetCenterX)/(eyeCenterLY-targetCenterY))
#   } else {
#     tAlphaL = -atan(abs(eyeCenterLX-targetCenterX)/(eyeCenterLY-targetCenterY))
#   }
#   
#   #noisyAlphaL = rnorm(1,tAlphaL, sdPer)
#   
#   if(eyeCenterRX > pupilRX) {
#     tAlphaR = atan(abs(eyeCenterRX-targetCenterX)/(eyeCenterRY-targetCenterY))
#   } else {
#     tAlphaR = -atan(abs(eyeCenterRX-targetCenterX)/(eyeCenterRY-targetCenterY))
#   }
#   
#   #noisyAlphaR = rnorm(1,tAlphaR, sdPer)
# 
#   bRDist = c()
#   
#   bLDist = c()
#   
#   for (x in 1:1920){
#     
#   if(eyeCenterLX > x) {
#     betaL = atan(abs(eyeCenterLX-x)/(eyeCenterLY-targetCenterY))
#   } else {
#     betaL = -atan(abs(eyeCenterLX-x)/(eyeCenterLY-targetCenterY))
#   }
#     betaLDist = dnorm(betaL,tAlphaL,sdTol, log = T)
#     
#   if(eyeCenterRX > x) {
#     betaR = atan(abs(eyeCenterRX-x)/(eyeCenterRY-targetCenterY))
#   } else {
#     betaR = -atan(abs(eyeCenterRX-x)/(eyeCenterRY-targetCenterY))
#   }
#     betaRDist = dnorm(betaR,tAlphaR,sdTol, log = T)
#     
#     #betaDist = betaLDist + betaRDist
#     
#     #bDist[x] <- exp(betaDist)
#     bRDist[x] <- betaRDist
#     
#     bLDist[x] <- betaLDist
#     
#   }
#   
#   xRDist = exp(bRDist-logSumExp(bRDist))
#   
#   xLDist = exp(bLDist-logSumExp(bLDist))
#   
#   xDist = xRDist * xLDist
#   
#   dist = xDist / sum(xDist)
#   
#   # click = sample(1:1920, prob = xDist, size = 1)-1
#   
#   return(dist)
# }
```

```{r headlight_dist}
# headlight_df <- data.frame(
#   coord = 1:1920,
#   sdTol = "0.175"
# )
# 
# for (i in c(960,1736)){
#   
#   probdist <-  get_dist(
#     sdTol = 0.175,
#     targetCenterX = i,
#     eyeCenterLY = 937.8,
#     eyeCenterLX = 872.12,
#     eyeCenterRY = 937.8,
#     eyeCenterRX = 1045.89,
#     targetCenterY = 127.5,
#     pupilLX = i-1,
#     pupilRX = i-1)
#   
#     colname = as.character(i)
#     
#     headlight_df[, colname] <- probdist
#   
#   }
# 
# headlight_df %>%
#   pivot_longer(names_to = "target_coordinate", values_to = "probability", cols = c(-coord, -sdTol)) %>%
#   filter(target_coordinate == "960") %>%
#   ggplot(aes(x = coord, y = probability, group = interaction(target_coordinate,sdTol))) +
#   geom_density(stat = "identity", color = "blue", fill = "blue", alpha = .5) +
#   ylim(0, 0.006) +
#   scale_x_continuous(limits = c(0, 1920), breaks = seq(0, 1920, 480)) +
#   guides(fill = F) + 
#   # geom_vline(aes(xintercept = as.numeric(target_coordinate)), lty = 2, alpha = .25) +
#   # labs(x = "Coordinate", y = "Probability") +
#   # scale_fill_colorblind(name = "Inference SD") +
#   theme_few() + 
#   theme(
#     panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
#     panel.grid.minor = element_blank(), 
#     panel.grid.major = element_blank(),
#     plot.background = element_rect(fill = "transparent",colour = NA)
# )
# 
# ggsave("../figures/gazemodel_headlight_dist_side.png", width = 15, height = 4, bg = "transparent")
# 
# ##  Standard normal dist
# dnorm_one_sd <- function(x){
#   norm_one_sd <- dnorm(x)
#   # Have NA values outside interval x in [-1, 1]:
#   norm_one_sd[x <= -1 | x >= 1] <- NA
#   return(norm_one_sd)
# }
# 
# ggplot(data.frame(x = c(-4, 4)), aes(x = x)) +
#   stat_function(fun = dnorm, color = "darkgrey", linewidth = 2) + 
#   stat_function(fun = dnorm_one_sd, geom = "area", fill = "blue", alpha = 0.5) +
#   # geom_vline(aes(xintercept = 0), lty = 2, color = "blue") +
#   theme_void()
# 
# ggsave("../figures/gazemodel_normal_dist.png", width = 10, height = 4)
```